---
title: "2022_11_4_UPDATED_Survival_Analysis_full_data"
output: html_document
date: '2022-10-04'
---


**UPDATED: ALL CLINICAL DATA**

# **Purpose:** 

Perform survival analysis for the full RIKEN cohort males vs. females and subset by etiology


# **Results:**

### Load libraries
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(survminer)
library(survival)
```

### Read in data
```{r}
metadata  <- read.table("~/Desktop/Research Projects/Liver Cancer Project/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",") #changing the name of the file
tumorAdjacentExp <- read.table("~/Desktop/Research Projects/Liver Cancer Project/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE) #changing the name of the file 
clinical_metadata <- read.csv("~/Desktop/Research Projects/Liver Cancer Project/Metadata/Full_RIKEN_clinical_data.csv")
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp))
```

### Match data
```{r}
clinical_metadata <- clinical_metadata[which(clinical_metadata$ID %in% metadata$ID),]
metadata <- metadata[which(metadata$ID %in% clinical_metadata$ID),]
clinical_metadata <- clinical_metadata[match(metadata$ID, clinical_metadata$ID),]

identical(metadata$ID, clinical_metadata$ID)
```

## Survival analysis all males vs. females: 
```{r}
for (i in 1:length(metadataSubset$donor_vital_status)){
  if (metadataSubset$donor_vital_status[i] == "deceased"){
    metadataSubset$donor_vital_status[i] = 1
  }
  if (metadataSubset$donor_vital_status[i] == "alive"){
    metadataSubset$donor_vital_status[i] =0
  }
}

ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata), pval=TRUE)

pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/All_data_survival.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata), pval=TRUE)
dev.off()
```

## Survival analysis subset males vs. females - HBV: 
```{r}
Total_metadata_HBV <- Total_metadata[which(Total_metadata$Virus.infection=="HBV"),]
  
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HBV), pval=TRUE)

pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/HBV_data_survival.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HBV), pval=TRUE)
dev.off()
```

## Survival analysis subset males vs. females - HCV: 
```{r}
Total_metadata_HCV <- Total_metadata[which(Total_metadata$Virus.infection=="HCV"),]
  
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HCV), pval=TRUE)

pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/HCV_data_survival.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HCV), pval=TRUE)
dev.off()
```

## Survival analysis subset males vs. females - Neither: 
```{r}
Total_metadata_NBNC <- Total_metadata[which(Total_metadata$Virus.infection=="NBNC"),]
  
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_NBNC), pval=TRUE)

pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/NBNC_data_survival.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_NBNC), pval=TRUE)
dev.off()
```

## Read in French Liver cancer data
```{r}
data_HCC <- as.data.frame(read.table("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/France Liver Cancer/donor.LICA-FR.tsv",sep="\t", header=TRUE))
data_HCC <- data_HCC[-which(data_HCC$donor_vital_status==""),]

for (i in 1:length(data_HCC$donor_vital_status)){
  if (data_HCC$donor_vital_status[i] == "deceased"){
    data_HCC$status[i] = 1
  }
  if (data_HCC$donor_vital_status[i] == "alive"){
    data_HCC$status[i] =0
  }
}

ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(status))~donor_sex, data=data_HCC), pval=TRUE)

pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/French_data_survival.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(status))~donor_sex, data=data_HCC), pval=TRUE)
dev.off()
```

**Interpretation:**

* Across a combination of all of the data, there is significantly worse prognosis for females than males

* In no subset is the difference significant

**Future steps:**

* Integrate into final figure

