---
title: "TSPY1AndXISTExpressionInvestigation"
author: "Annika Jorgensen"
date: "2023-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Purpose**
The purpose of this file is twofold. One purpose for this file see if the male tumor samples that show XIST expression graph (created in Sex_check_automation_removed_samples.Rmd) are the same samples that show TSPY1 expression. Another purpose is to see if the male tumor samples with XIST expression are also associated with low Y chromosome expression. 

***

## **Results**

### **Pseudo code **

* Read in expression data for the five y chromosome genes and XIST

* Subset data for all Male tumor samples.

* Find samples with TSPY1 expression
* Find samples with XIST expression. If any samples overlap with TSPY1 and XIST
  +  plot tumor expression of all y chromosome genes and XIST
  +  plot tumor adj expression as well 
* For all samples with XIST expression. See if samples have low y chromsome expression
  + See if they show up on any previous graphs made in past files 
  + If low expression make a note

```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)
library(tidyverse)

setwd('~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/GeneExpressionPlots')
```

```{r read in data}
#reading in tumor expression data for DDX3Y, EIF1AY, TXLNGY, KDM5D, TSPY1
Tumor_expressionData <- read.csv("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/GeneExpressionPlots/data/TumorExpressionData6genes.csv",row.names=1,header=TRUE, sep=",")
```

```{r subsetting by males and genes}
Tumor_expressionData<- t(Tumor_expressionData)
Tumor_expressionData<- as.data.frame(Tumor_expressionData)

#subsetting by male tumor samples
MaleTumor_expressionData<- Tumor_expressionData[grepl("XY",rownames(Tumor_expressionData),fixed=TRUE),]
MaleTumor_expressionData<- t(MaleTumor_expressionData)
MaleTumor_expressionData<- as.data.frame(MaleTumor_expressionData)
```

**Samples with XIST Expression**
* Sample 079
  + Contains TSPY1 expression (low amounts)
  
* Sample 027
  + No TSPY1 expression
```{r plot XIST expression males}
#sample 027 expression 
dataFiltered<-  as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,13],Names=c("027","027","027","027","027","027")))
rownames(dataFiltered)<-1:6

#samples 079 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,38],Names=c("079","079","079","079","079","079")))
rownames(newRow)<-1:6

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 151 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,66],Names=c("151","151","151","151","151","151")))
rownames(newRow)<-1:6

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 175 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,75],Names=c("175","175","175","175","175","175")))
rownames(newRow)<-1:6

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 209 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,85],Names=c("209","209","209","209","209","209")))
rownames(newRow)<-1:6

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 289 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,108],Names=c("289","289","289","289","289","289")))
rownames(newRow)<-1:6

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)
```

```{r plotting expression}
#changing Expression values to be numeric 
dataFiltered$Expression<- as.numeric(dataFiltered$Expression)

#plotting all samples with FPKM under 500
expressionPlot<- ggplot(data=dataFiltered, aes(x=genes,y=Expression,group=Names)) +  geom_line(aes(color=Names))+
  geom_point(aes(color=Names)) + ggtitle("Male Tumor Samples w/ High XIST expression") + ylab("FPKM") +xlab("")+ labs(color = "Samples") #+ ylim(0,50)
expressionPlot
```
Plotted six samples with High XIST expression. Out of the six samples, *three samples had zero or low TSPY1 expression (sample 175, and 027 had zero and sample 079 had low (under 100 FPKM)).* 

```{r plot without XIST and TSPY1}

Tumor_expressionData<- subset(Tumor_expressionData, select = -c(XIST,TSPY1))

#subsetting by male tumor samples
MaleTumor_expressionData<- Tumor_expressionData[grepl("XY",rownames(Tumor_expressionData),fixed=TRUE),]
MaleTumor_expressionData<- t(MaleTumor_expressionData)
MaleTumor_expressionData<- as.data.frame(MaleTumor_expressionData)
```

```{r dataframe without TSPY1 and XIST}
#sample 027 expression 
dataFiltered<-  as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,13],Names=c("027","027","027","027")))
rownames(dataFiltered)<-1:4

#samples 079 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,38],Names=c("079","079","079","079")))
rownames(newRow)<-1:4

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 151 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,66],Names=c("151","151","151","151")))
rownames(newRow)<-1:4

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 175 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,75],Names=c("175","175","175","175")))
rownames(newRow)<-1:4

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 209 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,85],Names=c("209","209","209","209")))
rownames(newRow)<-1:4

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)

#samples 289 expression 
newRow<- as.data.frame(cbind(genes=rownames(MaleTumor_expressionData),Expression=MaleTumor_expressionData[,108],Names=c("289","289","289","289")))
rownames(newRow)<-1:4

#combing expression into data frame
dataFiltered<- rbind(dataFiltered,newRow)
```

```{r plotting expression without TSPY1 and XIST}
#changing Expression values to be numeric 
dataFiltered$Expression<- as.numeric(dataFiltered$Expression)

#plotting all samples with FPKM under 500
expressionPlot<- ggplot(data=dataFiltered, aes(x=genes,y=Expression,group=Names)) +  geom_line(aes(color=Names))+
  geom_point(aes(color=Names)) + ggtitle("Male Tumor Samples w/ High XIST expression") + ylab("FPKM") +xlab("")+ labs(color = "Samples") #+ ylim(0,50)
expressionPlot
```

It does not appear that male tumor samples with high XIST expression have low Y chromosome expression.