---
title: "TumorAdjSamplesGraphNoTSPY1Gene"
author: "Annika Jorgensen"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Purpose**
This Rmarkdown will plot FPKM expression of DDX3Y, EIF1AY, TXLNGY, KDM5D separated by male tumor adj samples. This is a second iteration of XdegeneratFPKMespression.Rmd where tumor samples FPKM expression were plotted but not separated by sex and the gene TSPY1 was added. Sex is separated out because these genes are y-linked. TSPY1 is not included because there is no expression in most samples.

Refer to the TumorSamplesGraphNoTSPY1Gene.Rmd markdown file for how to output the male tumor sample gene expression graphs.

## **Results**
 Two graphs were created. The first showed the gene expression of all samples with expression under **100 FPKM**. The second showed the gene expression of all samples with expression under **500 FPKM**.
 
```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)
library(tidyverse)

setwd('~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/GeneExpressionPlots')
```

**Read in data**

```{r reading in data}
#reading in tumor expression data for DDX3Y, EIF1AY, TXLNGY, KDM5D, TSPY1
TumorAdj_expressionData <- read.csv("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/GeneExpressionPlots/data/TumorAdjExpressionData5genes.csv",row.names=1,header=TRUE, sep=",")

#removing expression data for TSPY1
TumorAdj_expressionData<- TumorAdj_expressionData[-5,]

#creating a vector for the x axis of our plot
geneXaxis<- 1:4

#reading in tumor expression data for DDX3Y, EIF1AY, TXLNGY, KDM5D, TSPY1
Tumor_expressionData <- read.csv("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/GeneExpressionPlots/data/TumorExpressionData5genes.csv",row.names=1,header=TRUE, sep=",")

#removing expression data for TSPY1
Tumor_expressionData<- Tumor_expressionData[-5,]

#creating a vector for the x axis of our plot
geneXaxis<- 1:4
```

Making a data frame with all male tumor samples
```{r MaleTumorExpress}
TumorAdj_expressionData<- t(TumorAdj_expressionData)
TumorAdj_expressionData<- as.data.frame(TumorAdj_expressionData)

MaleAdj_expressionData<- TumorAdj_expressionData[grepl("XY",rownames(TumorAdj_expressionData),fixed=TRUE),]
MaleAdj_expressionData<- t(MaleAdj_expressionData)
MaleAdj_expressionData<-as.data.frame(MaleAdj_expressionData)

TumorAdj_expressionData<- t(TumorAdj_expressionData)
TumorAdj_expressionData<- as.data.frame(TumorAdj_expressionData)
```

**Tumor, Tumor Adj Plot for sample 003**
Plot shows Tumor Tumor Adj expression for sample 003. Tried to plot sample 207 as well but there is no Tumor Adjacent Sample for it
```{r paired data frame}
#creating data frame with sample 3 TumorAdj expression data
dataFiltered<-  as.data.frame(cbind(genes=rownames(MaleAdj_expressionData),Expression=MaleAdj_expressionData[,2],Names=c("Normal","Normal","Normal","Normal")))
rownames(dataFiltered)<-1:4

#creating data frame with sample 3 Tumor expression
newRow<- as.data.frame(cbind(genes=rownames(Tumor_expressionData),Expression=Tumor_expressionData[,2],Names=c("Tumor","Tumor","Tumor","Tumor")))
rownames(newRow)<-1:4

#combining sample 3 Tumor with sample 3 TumorAdj
dataFiltered<- rbind(dataFiltered,newRow)
```

```{r paired output plot}
#changing Expression values to be numeric 
dataFiltered$Expression<- as.numeric(dataFiltered$Expression)

#plotting all samples with FPKM under 100
expressionPlot<- ggplot(data=dataFiltered, aes(x=genes,y=Expression,group=Names)) +  geom_line(aes(color=Names))+
  geom_point(aes(color=Names))+ ggtitle("Sample 3 Gene Expression") + ylab("FPKM") +xlab("")+ labs(color = "Samples")
expressionPlot

pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/FPKM_sampleNamesPlots/Sample003GeneExpression.pdf", width=12, height=12)
#plotting all samples with FPKM under 100
expressionPlot<- ggplot(data=dataFiltered, aes(x=genes,y=Expression,group=Names)) +  geom_line(aes(color=Names))+
  geom_point(aes(color=Names))+ ggtitle("Sample 3 Gene Expression") + ylab("FPKM") +xlab("")+ labs(color = "Samples")
expressionPlot
dev.off()
```
**Tumor, Tumor Adj Plot for sample 169**

Plotting Tumor Tumor-Adj gene expression because it is the only sample with under 500FPKM for all four genes.
```{r under 500FPKM}
#creating data frame with sample 169 TumorAdj expression data
dataFiltered<-  as.data.frame(cbind(genes=rownames(MaleAdj_expressionData),Expression=MaleAdj_expressionData[,65],Names=c("Normal","Normal","Normal","Normal")))
rownames(dataFiltered)<-1:4

#creating data frame with sample 169 Tumor expression
newRow<- as.data.frame(cbind(genes=rownames(Tumor_expressionData),Expression=Tumor_expressionData[,98],Names=c("Tumor","Tumor","Tumor","Tumor")))
rownames(newRow)<-1:4

#combining sample 169 Tumor with sample 169 TumorAdj
dataFiltered<- rbind(dataFiltered,newRow)
```

```{r paired output plot 169}
#changing Expression values to be numeric 
dataFiltered$Expression<- as.numeric(dataFiltered$Expression)

#plotting gene expression sample 169
expressionPlot<- ggplot(data=dataFiltered, aes(x=genes,y=Expression,group=Names)) +  geom_line(aes(color=Names))+
  geom_point(aes(color=Names))+ ggtitle("Sample 169 Gene Expression") + ylab("FPKM") +xlab("")+ labs(color = "Samples")
expressionPlot

pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/FPKM_sampleNamesPlots/Sample169GeneExpression.pdf", width=12, height=12)
#plotting gene expression sample 169
expressionPlot<- ggplot(data=dataFiltered, aes(x=genes,y=Expression,group=Names)) +  geom_line(aes(color=Names))+
  geom_point(aes(color=Names))+ ggtitle("Sample 169 Gene Expression") + ylab("FPKM") +xlab("")+ labs(color = "Samples")
expressionPlot
dev.off()
```