---
title: "Medical Covariate Analysis"
author: "Annika Jorgensen"
date: "2023-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)
library(BSDA)
```

```{r  Reading in data}
#reading in Full meta dataset
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/Full_RIKEN_clinical_data.csv", row.names=1,header=TRUE, sep=",")

#reading in meta dataset
metadataSmall <- read.csv("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/MedicalCovariateAnalysis/data/metadataSubset.csv", row.names=1)
```

```{r subsetting data}
#subsetting full metadata for samples used in analysis
metadataAnalysis <- metadata[rownames(metadata) %in% metadataSmall$ID,]

#subsetting male samples 
Male <- metadataAnalysis[which(metadataAnalysis$Gender== "M"),]

#subsetting female samples 
Female <- metadataAnalysis[which(metadataAnalysis$Gender== "F"),]
```

***

### **F and T test age of males and females**

```{r Age difference data}
#subsetting age of male samples
MaleAge<- Male$Age

#subsetting age of female samples
FemaleAge<- Female$Age

#f test alpha=0.05
var.test(MaleAge, FemaleAge) #Fail to reject null no evidence variance different

#variances equal
```

No evidence that variances are equal. So, we will conduct a t test with equal variance

T test conducted on ages 
```{r t test age }
#t test alpha=0.05
t.test(MaleAge,FemaleAge,var.equal=TRUE) #Fail to reject null mean age of males and females not significantly different
```
The ages of male samples compared to female samples is not significantly different. It should be noted that the mean age of males is 65.23737 and the mean age of females is 68.76250

CONCLUSION: With 95% confidence we can say that the mean age of male patients compared to female patients is not significantly different.

***

### **F and T test tumor size in mm**
```{r f test tumor size}
#subsetting by tumor size in female samples 
FemaleSize <- Female$Tumor.size..mm.

#subsetting by tumor size in male samples
MaleSize<- Male$Tumor.size..mm.

#f test alpha=0.05
var.test(MaleSize, FemaleSize) #reject null variances significantly different 

#variances unequal
```

Variances in male tumor size in mm is significantly different than variance in female tumor size. So, we will conduce a t test with unequal variance.
```{r t test tumor size}
#t test tumor size alpha=0.05
t.test(MaleSize,FemaleSize) #Do not reject null no evidence tumor size is different

#no significant difference
```

The tumor size is not significantly different between males and females. 

CONCLUSION: With 95% confidence we conclude that the mean tumor size (in mm) is not significantly different between males and females.

***

### **F and T test overall survival month**
```{r f test overall survival month}
#subsetting overall survival month from male sample
MaleSurvivalMonth<- Male$Overall.survival..month.

#subsetting overall survival month from female sample
FemaleSurvivalMonth<- Female$Overall.survival..month.

#f test overall survival month alpha=0.05
var.test(MaleSurvivalMonth, FemaleSurvivalMonth) #fail to reject null no evidence variances are different

#variances are equal
```

```{r t test overall survival month}
#t test overall survival month alpha=0.05
t.test(MaleSurvivalMonth, FemaleSurvivalMonth, var.equal = TRUE) #Reject null survival month significantly different

#significant difference
```

Significant difference between the mean survival month of males compared to females

CONCLUSION: With 95% confidence we conclude that the mean survival month of males is significantly different than the mean survival month of females. 

Also, let it be known that the mean survival month of females is five month shorter than the mean survival month of males. This is inconsistent with the literature.

***

### **Test of Proportions Smoking Males and Females** 
```{r smoking z test}
#proportion of male smokers
MaleSmokers<-sum(Male$Smoking)
MaleSmokers


#proportion of female smokers
FemaleSmokers<- sum(Female$Smoking)
FemaleSmokers

```

```{r two sample z test}
# alpha=0.05
prop.test(x=c(MaleSmokers,FemaleSmokers), n= c(108,42), p = NULL, alternative = "two.sided", correct = TRUE) #reject null hypothesis that there is no difference between the number of men and women that smoke

#the number of men and women that smoke is significantly different
```
CONCLUSION: We are 95% confident that the number of men and women that smoke is significantly different. It should be noted that out 83 men smoke out of a sample of 108 and 4 women smoke out of a sample of 42. 

***

## **CHI-squared Alcohol**

```{r Alcohol}
#male alcohol intake separated by severity 
Male0<- nrow(Male[Male$Alcohol.intakee== "0", ])
Male1<- nrow(Male[Male$Alcohol.intakee== "1", ])
Male2<- nrow(Male[Male$Alcohol.intakee== "2", ])
Male3<- nrow(Male[Male$Alcohol.intakee== "3", ])

#female alcohol
Female0<- nrow(Female[Female$Alcohol.intakee== "0", ])
Female1<- nrow(Female[Female$Alcohol.intakee== "1", ])
Female2<- nrow(Female[Female$Alcohol.intakee== "2", ])
Female3<- nrow(Female[Female$Alcohol.intakee== "3", ])

#counts of alcohol intake

AlcoholIntake<- as.data.frame(cbind(Males= c(Male0,Male1,Male2,Male3), Females= c(Female0, Female1, Female2,Female3)))
rownames(AlcoholIntake)<- c(0,1,2,3)

#sums of rows in Alcoholintake
AlcoholTotalCounts<-c(63,25,42,20)

#add sums of rows to AlcoholIntake
AlcoholIntake<- cbind(AlcoholIntake,AlcoholTotalCounts)

#sums of columns in AlcoholIntake
AlcoholTotalCounts<-c(108,42,150)

#add sums of column to AlcoholIntake
AlcoholIntake<- rbind(AlcoholIntake,AlcoholTotalCounts)
rownames(AlcoholIntake)<- c("0","1","2","3","Totals")

#creating expected counts
ExpectedAlcoholIntake<- as.data.frame(cbind(Males=c((AlcoholIntake[5,1]*AlcoholIntake[1,3])/AlcoholIntake[5,3],(AlcoholIntake[5,1]*AlcoholIntake[2,3])/AlcoholIntake[5,3],(AlcoholIntake[5,1]*AlcoholIntake[3,3])/AlcoholIntake[5,3],(AlcoholIntake[5,1]*AlcoholIntake[4,3])/AlcoholIntake[5,3]),Females=c((AlcoholIntake[5,2]*AlcoholIntake[1,3])/AlcoholIntake[5,3],(AlcoholIntake[2,3]*AlcoholIntake[5,2])/AlcoholIntake[5,3],(AlcoholIntake[5,2]*AlcoholIntake[3,3])/AlcoholIntake[5,3],(AlcoholIntake[5,2]*AlcoholIntake[4,3])/AlcoholIntake[5,3])))
```

Critical statistic value taken from [here](https://www.math.arizona.edu/~jwatkins/chi-square-table.pdf)
```{r CHI Alcohol}
#calculating chi square statistic 
((AlcoholIntake[2,1]-ExpectedAlcoholIntake[2,1])^2/ExpectedAlcoholIntake[2,1]) +
((AlcoholIntake[3,1]-ExpectedAlcoholIntake[3,1])^2/ExpectedAlcoholIntake[3,1]) + ((AlcoholIntake[4,1]-ExpectedAlcoholIntake[4,1])^2/ExpectedAlcoholIntake[4,1]) + ((AlcoholIntake[2,2]-ExpectedAlcoholIntake[2,2])^2/ExpectedAlcoholIntake[2,2]) + ((AlcoholIntake[3,2]-ExpectedAlcoholIntake[3,2])^2/ExpectedAlcoholIntake[3,2]) + ((AlcoholIntake[4,2]-ExpectedAlcoholIntake[4,2])^2/ExpectedAlcoholIntake[4,2])

#the statistic from the chi distribution table yields a critical statistic of 7.815 based on a alpha 0.05 and 3 degrees of freedom which is less than 13.968. Because the critical statistic is less than our chi square value we can reject the null hypothesis that the proportion of males and females who intake alcohol are the same.

#proportions different
```

CONCLUSION: We are 95 percent confident that the proportion of males and females who intake alcohol are not the same. 

Listed below is the table show the counts of males and females who intake alcohol. The row names are the levels of alcohol severity, where zero being patients who do not intake alcohol (therefore not included in the chi square test).

```{r Table}
AlcoholIntake
```

*** 

## **Liver Fibrosis CHI-square**
```{r Liver Fibrosis}
#male fibrosis separated by severity 
Male0F<- nrow(Male[Male$Liver.fibrosisc== "0", ])
Male1F<- nrow(Male[Male$Liver.fibrosisc== "1", ])
Male2F<- nrow(Male[Male$Liver.fibrosisc== "2", ])
Male3F<- nrow(Male[Male$Liver.fibrosisc== "3", ])
Male4F<- nrow(Male[Male$Liver.fibrosisc== "4", ])

#female fibrosis
Female0F<- nrow(Female[Female$Liver.fibrosisc== "0", ])
Female1F<- nrow(Female[Female$Liver.fibrosisc== "1", ])
Female2F<- nrow(Female[Female$Liver.fibrosisc== "2", ])
Female3F<- nrow(Female[Female$Liver.fibrosisc== "3", ])
Female4F<- nrow(Female[Female$Liver.fibrosisc== "4", ])

#counts of liver fibrosis

LiverFibrosis<- as.data.frame(cbind(Males= c(Male0F,Male1F,Male2F,Male3F,Male4F), Females= c(Female0F, Female1F, Female2F,Female3F,Female4F)))
rownames(LiverFibrosis)<- c(0,1,2,3,4)

#sums of rows in LiverFibrosis
LiverFibrosisCounts<-c(5,16,30,40,59)

#add sums of rows to LiverFibrosis
LiverFibrosis<- cbind(LiverFibrosis,LiverFibrosisCounts)

#sums of columns in LiverFibrosis
LiverTotalCounts<-c(108,42,150)

#add sums of column to LiverFibrosis
LiverFibrosis<- rbind(LiverFibrosis,LiverTotalCounts)
rownames(LiverFibrosis)<- c("0","1","2","3","4","Totals")
```

```{r LiverExpectedCounts}
#creating expected counts
ExpectedLiverFibrosis<- as.data.frame(cbind(Males=c((LiverFibrosis[6,1]*LiverFibrosis[1,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[2,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[3,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[4,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[2,3])/ LiverFibrosis[5,3]),Females=c((LiverFibrosis[6,2]*LiverFibrosis[1,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,2]*LiverFibrosis[2,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[3,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[4,3])/ LiverFibrosis[6,3],(LiverFibrosis[6,1]*LiverFibrosis[5,3])/ LiverFibrosis[6,3])))
```

```{r CHI square stat Liver Fibrosis}
#calculating Chi square statistic
(LiverFibrosis[2,1]-ExpectedLiverFibrosis[2,1])^2/ExpectedLiverFibrosis[2,1] + (LiverFibrosis[3,1]-ExpectedLiverFibrosis[3,1])^2/ExpectedLiverFibrosis[3,1] + (LiverFibrosis[4,1]-ExpectedLiverFibrosis[4,1])^2/ExpectedLiverFibrosis[4,1] + (LiverFibrosis[5,1]-ExpectedLiverFibrosis[5,1])^2/ExpectedLiverFibrosis[5,1] + (LiverFibrosis[2,2]-ExpectedLiverFibrosis[2,2])^2/ExpectedLiverFibrosis[2,2] + (LiverFibrosis[3,2]-ExpectedLiverFibrosis[3,2])^2/ExpectedLiverFibrosis[3,2] + (LiverFibrosis[4,2]-ExpectedLiverFibrosis[4,2])^2/ExpectedLiverFibrosis[4,2] + (LiverFibrosis[5,2]-ExpectedLiverFibrosis[5,2])^2/ExpectedLiverFibrosis[5,2]

#the statistic from the chi distribution table yields a critical statistic of 9.488 based on a alpha 0.05 and 4 degrees of freedom which is less than 36.966. Because the critical statistic is less than our chi square value we can reject the null hypothesis that the proportion of males and females who have liver fibrosis are the same.

#proportions different
```


CONCLUSION: We are 95 percent confident that the proportion of males and females who have liver fibrosis are not the same. 

Listed below is the table show the counts of males and females who have liver fibrosis. The row names are the levels of liver fibrosis severity, where zero being patients who do not have liver fibrosis (therefore not included in the chi square test).

```{r Liver Fibrosis Table}
#outputting table
LiverFibrosis
```

***

## **Tumor Stage CHI-square**
Creating a data frame of counts of males and females with each stage
```{r counts}
#male stage separated by severity 
Male1st<- nrow(Male[Male$Ta== "1", ])
Male2st<- nrow(Male[Male$Ta== "2", ])
Male3st<- nrow(Male[Male$Ta== "3", ])
Male4st<- nrow(Male[Male$Ta== "4", ])

#female stage separated by severity
Female1st<- nrow(Female[Female$Liver.fibrosisc== "1", ])
Female2st<- nrow(Female[Female$Liver.fibrosisc== "2", ])
Female3st<- nrow(Female[Female$Liver.fibrosisc== "3", ])
Female4st<- nrow(Female[Female$Liver.fibrosisc== "4", ])

#counts of tumor stage
TumorStage<- as.data.frame(cbind(Males= c(Male1st,Male2st,Male3st,Male4st), Females= c(Female1st, Female2st, Female3st,Female4st)))
rownames(TumorStage)<- c(1,2,3,4)

#sums of rows in TumorStage
TumorStageCounts<-c(20,53,50,27)

#add sums of rows to TumorSTage
TumorStage<- cbind(TumorStage,TumorStageCounts)

#sums of columns in TumorStage
StageTotalCounts<-c(108,42,150)

#add sums of column to TumorStage
TumorStage<- rbind(TumorStage,StageTotalCounts)
rownames(TumorStage)<- c("1","2","3","4","Totals")
```
And important note here is that 20 out of 42 female samples are in stage four

```{r StageExpected Counts}
#calculating expected counts
ExpectedStageCounts<-  as.data.frame(cbind(Males=c((TumorStage[1,3]*TumorStage[5,1])/TumorStage[5,3],(TumorStage[2,3]*TumorStage[5,1])/TumorStage[5,3],(TumorStage[3,3]*TumorStage[5,1])/TumorStage[5,3],(TumorStage[4,3]*TumorStage[5,1])/TumorStage[5,3]), Females=c((TumorStage[1,3]*TumorStage[5,2])/TumorStage[5,3],(TumorStage[2,3]*TumorStage[5,2])/TumorStage[5,3],(TumorStage[3,3]*TumorStage[5,2])/TumorStage[5,3],(TumorStage[4,3]*TumorStage[5,2])/TumorStage[5,3])))
```

```{r Chi squared statistic tumor stage}
#calculating Chi square statistic
(TumorStage[1,1]-ExpectedStageCounts[1,1])^2/ExpectedStageCounts[1,1] + (TumorStage[2,1]-ExpectedStageCounts[2,1])^2/ExpectedStageCounts[2,1] +
(TumorStage[3,1]-ExpectedStageCounts[3,1])^2/ExpectedStageCounts[3,1] +
(TumorStage[4,1]-ExpectedStageCounts[4,1])^2/ExpectedStageCounts[4,1] +
(TumorStage[1,2]-ExpectedStageCounts[1,2])^2/ExpectedStageCounts[1,2] +
(TumorStage[2,2]-ExpectedStageCounts[2,2])^2/ExpectedStageCounts[2,2] +
(TumorStage[3,2]-ExpectedStageCounts[3,2])^2/ExpectedStageCounts[3,2] +
(TumorStage[4,2]-ExpectedStageCounts[4,2])^2/ExpectedStageCounts[4,2]

#the statistic from the chi distribution table yields a critical statistic of 7.815 based on a alpha 0.05 and 3 degrees of freedom which is less than 39.169. Because the critical statistic is less than our chi square value we can reject the null hypothesis that the proportion of males and females with each tumor stage are the same.

#proportions different
```

CONCLUSION: We are 95 percent confident that the proportion of males and females with each tumor stage are not the same.

An important note to make here is almost have of the female samples (20 out of 42) have stage four. This may be contributing to the lower survival month of females.

Listed below is the table show the counts of males and females wth each tumor stage. The row names are the levels of tumor stage.

```{r tumor stage table}
TumorStage
```

***

## **CHI square Edmondson Grade**
Completing CHI square test on Edmondson Grade. Values denotes as 2~3 and 1~2 were changed to 2.5 and 1.5 respectively 
```{r Edmondson Grade}
#changing cells to work with CHI square test
#Male["Edmondson.grade"][Male["Edmondson.grade"] == "2~3"] <- "2.5"
#Male["Edmondson.grade"][Male["Edmondson.grade"] == "1~2"] <- "1.5"
Male[is.na(Male)] <-0

#Female["Edmondson.grade"][Female["Edmondson.grade"] == "2~3"] <- "2.5"
#Female["Edmondson.grade"][Female["Edmondson.grade"] == "1~2"] <- "1.5"
Female[is.na(Female)] <-0

#male grade separated by severity 
MaleEg0<- nrow(Male[Male$Edmondson.grade=="0", ])
MaleEg1<- nrow(Male[Male$Edmondson.grade=="1", ])
MaleEg1.5<- nrow(Male[Male$Edmondson.grade=="1~2", ])
MaleEg2<- nrow(Male[Male$Edmondson.grade=="2", ])
MaleEg2.5<- nrow(Male[Male$Edmondson.grade=="2~3", ])
MaleEg3<- nrow(Male[Male$Edmondson.grade=="3", ])
MaleEg4<- nrow(Male[Male$Edmondson.grade=="4", ])

#female grade separated by severity 
FemaleEg0<- nrow(Female[Female$Edmondson.grade=="0", ])
FemaleEg1<- nrow(Female[Female$Edmondson.grade=="1", ])
FemaleEg1.5<- nrow(Female[Female$Edmondson.grade=="1~2", ])
FemaleEg2<- nrow(Female[Female$Edmondson.grade=="2", ])
FemaleEg2.5<- nrow(Female[Female$Edmondson.grade=="2~3", ])
FemaleEg3<- nrow(Female[Female$Edmondson.grade=="3", ])
FemaleEg4<- nrow(Female[Female$Edmondson.grade=="4", ])

#counts of Edmonson grade
EdmonsonGrade<- as.data.frame(cbind(Males= c(MaleEg0,MaleEg1,MaleEg1.5,MaleEg2,MaleEg2.5,MaleEg3,MaleEg4), Females= c(FemaleEg0,FemaleEg1, FemaleEg1.5, FemaleEg2,FemaleEg2.5,FemaleEg3,FemaleEg4)))
rownames(EdmonsonGrade)<- c(0,1,1.5,2,2.5,3,4)

#sums of rows in EdmonsonGrade
EdmonsonGradeCounts<-c(1,14,8,88,23,16,1)

#add sums of rows to EdmonsonGrade
EdmonsonGrade<- cbind(EdmonsonGrade,EdmonsonGradeCounts)

#sums of columns in EdmonsonGrade
GradeTotalCounts<-c(108,42,150)

#add sums of column to EdmonsonGrade
EdmonsonGrade<- rbind(EdmonsonGrade,GradeTotalCounts)
rownames(EdmonsonGrade)<- c("0","1","1.5","2","2.5","3","4","Totals")
```

```{r Expected Counts Edmondson Grade}
#calculating expected counts
ExpectedEdmonson<- as.data.frame(cbind(Males=c((EdmonsonGrade[1,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3],(EdmonsonGrade[2,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3],(EdmonsonGrade[3,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3],(EdmonsonGrade[4,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3],(EdmonsonGrade[5,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3],(EdmonsonGrade[6,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3],(EdmonsonGrade[7,3]*EdmonsonGrade[8,1])/EdmonsonGrade[8,3]), Females=c((EdmonsonGrade[1,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3],(EdmonsonGrade[2,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3],(EdmonsonGrade[3,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3],(EdmonsonGrade[4,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3],(EdmonsonGrade[5,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3],(EdmonsonGrade[6,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3],(EdmonsonGrade[7,3]*EdmonsonGrade[8,2])/EdmonsonGrade[8,3])))
```

```{r Chi squared statistic Edmondson grade}
#calculating CHI squared statistic
(EdmonsonGrade[1,1]-ExpectedEdmonson[1,1])^2/ExpectedEdmonson[1,1] +
(EdmonsonGrade[2,1]-ExpectedEdmonson[2,1])^2/ExpectedEdmonson[2,1] +
(EdmonsonGrade[3,1]-ExpectedEdmonson[3,1])^2/ExpectedEdmonson[3,1] +
(EdmonsonGrade[4,1]-ExpectedEdmonson[4,1])^2/ExpectedEdmonson[4,1] +
(EdmonsonGrade[5,1]-ExpectedEdmonson[5,1])^2/ExpectedEdmonson[5,1] +
(EdmonsonGrade[6,1]-ExpectedEdmonson[6,1])^2/ExpectedEdmonson[6,1] +
(EdmonsonGrade[7,1]-ExpectedEdmonson[7,1])^2/ExpectedEdmonson[7,1] +
(EdmonsonGrade[1,2]-ExpectedEdmonson[1,2])^2/ExpectedEdmonson[1,2] +
(EdmonsonGrade[2,2]-ExpectedEdmonson[2,2])^2/ExpectedEdmonson[2,2] +
(EdmonsonGrade[3,2]-ExpectedEdmonson[3,2])^2/ExpectedEdmonson[3,2] +
(EdmonsonGrade[4,2]-ExpectedEdmonson[4,2])^2/ExpectedEdmonson[4,2] +
(EdmonsonGrade[5,2]-ExpectedEdmonson[5,2])^2/ExpectedEdmonson[5,2] +
(EdmonsonGrade[6,2]-ExpectedEdmonson[6,2])^2/ExpectedEdmonson[6,2] +
(EdmonsonGrade[7,2]-ExpectedEdmonson[7,2])^2/ExpectedEdmonson[7,2] 

#the statistic from the chi distribution table yields a critical statistic of 12.592 based on a alpha 0.05 and 6 degrees of freedom which is greater than 2.647. Because the critical statistic is less than our chi square value we cannot reject the null hypothesis that the proportion of males and females with each Edmondson grade are the same.

#proportions the same
```

CONCLUSION: We are 95 percent confident that the proportion of males and females with each Edmondson Grade are the same.

Listed below is the table show the counts of males and females with each Edmondson grade The row names are the levels of Edmondson Grade

```{r Edmondson grade table}
EdmonsonGrade
```
***

## **CHI square Prognosis**
```{r Prognosis}

#male prognosis separated by severity?
MaleProg0<- nrow(Male[Male$Prognosisf=="0",])
MaleProg1<- nrow(Male[Male$Prognosisf=="1",])
MaleProg2<- nrow(Male[Male$Prognosisf=="2",])
MaleProg3<- nrow(Male[Male$Prognosisf=="3",])

#female prognosis separated by severity?
FemaleProg0<- nrow(Female[Female$Prognosisf=="0",])
FemaleProg1<- nrow(Female[Female$Prognosisf=="1",])
FemaleProg2<- nrow(Female[Female$Prognosisf=="2",])
FemaleProg3<- nrow(Female[Female$Prognosisf=="3",])

#counts of Prognosis
Prognosis<- as.data.frame(cbind(Males= c(MaleProg0,MaleProg1,MaleProg2,MaleProg3), Females= c(FemaleProg0, FemaleProg1,FemaleProg2,FemaleProg3)))
rownames(Prognosis)<- c(0,1,2,3)

#sums of rows in Prognosis
PrognosisCounts<-c(124,21,3,2)

#add sums of rows to Prognosis
Prognosis<- cbind(Prognosis,PrognosisCounts)

#sums of columns in Prognosis
PrognosisTotalCounts<-c(108,42,150)

#add sums of column to Prognosis
Prognosis<- rbind(Prognosis,PrognosisTotalCounts)
rownames(Prognosis)<- c("0","1","2","3","Totals")
```

```{r Expected counts Prognosis}
ExpectedPrognosis<- as.data.frame(cbind(Males=c((Prognosis[1,3]*Prognosis[5,1])/Prognosis[5,3],(Prognosis[2,3]*Prognosis[5,1])/Prognosis[5,3],(Prognosis[3,3]*Prognosis[5,1])/Prognosis[5,3],(Prognosis[4,3]*Prognosis[5,1])/Prognosis[5,3]),Females=c((Prognosis[1,3]*Prognosis[5,2])/Prognosis[5,3],(Prognosis[2,3]*Prognosis[5,2])/Prognosis[5,3],(Prognosis[3,3]*Prognosis[5,2])/Prognosis[5,3],(Prognosis[4,3]*Prognosis[5,2])/Prognosis[5,3])))
```

```{r CHI statistic Prognosis}
(Prognosis[1,1]-ExpectedPrognosis[1,1])^2/ExpectedPrognosis[1,1] +
(Prognosis[2,1]-ExpectedPrognosis[2,1])^2/ExpectedPrognosis[2,1] +
(Prognosis[3,1]-ExpectedPrognosis[3,1])^2/ExpectedPrognosis[3,1] +
(Prognosis[4,1]-ExpectedPrognosis[4,1])^2/ExpectedPrognosis[4,1] +
(Prognosis[1,2]-ExpectedPrognosis[1,2])^2/ExpectedPrognosis[1,2] +
(Prognosis[2,2]-ExpectedPrognosis[2,2])^2/ExpectedPrognosis[2,2] +
(Prognosis[3,2]-ExpectedPrognosis[3,2])^2/ExpectedPrognosis[3,2] +
(Prognosis[4,2]-ExpectedPrognosis[4,2])^2/ExpectedPrognosis[3,2]

#the statistic from the chi distribution table yields a critical statistic of 5.991 based on a alpha 0.05 and 2 degrees of freedom which is greater than 2.647. Because the critical statistic is less than our chi square value we cannot reject the null hypothesis that the proportion of males and females with each Prognosis are the same.

#proportions the same
```

CONCLUSION: We are 95 percent confident that the proportion of males and females with each Prognosis are the same.

Listed below is the table show the counts of males and females with each Prognosis value The row names are the levels of Prognosis

```{r Prognosis table}
Prognosis
```
***
## **Test of Proportions Library Type** 
```{r proportions strandedness}

#subsetting male samples small metadata
MaleSm <- metadataSmall[which(metadataSmall$Gender== "M"),]

#subsetting female samples small metadata
FemaleSm <- metadataSmall[which(metadataSmall$Gender== "F"),]

#proportion of stranded male samples
MaleStranded<- nrow(MaleSm[MaleSm$strandedness=="stranded",])
MaleStranded
#proportion of stranded female samples
FemaleStranded<- nrow(FemaleSm[FemaleSm$strandedness=="stranded",])
FemaleStranded

# two sample z test alpha=0.05
prop.test(x=c(MaleStranded,FemaleStranded), n= c(198,80), p = NULL, alternative = "two.sided", correct = TRUE)

#not significantly different
```
CONCLUSION: We are 95% confident that the proportion of males and females with stranded DNA is not significantly different

```{r proportions unstrandedness }

#proportion of stranded male samples
MaleUnstranded<- nrow(MaleSm[MaleSm$strandedness=="unstranded",])
MaleUnstranded
#proportion of stranded female samples
FemaleUnstranded<- nrow(FemaleSm[FemaleSm$strandedness=="unstranded",])
FemaleUnstranded

# two sample z test alpha=0.05
prop.test(x=c(MaleUnstranded,FemaleUnstranded), n= c(198,80), p = NULL, alternative = "two.sided", correct = TRUE)

#Not significantly different
```
CONCLUSION: We are 95% confident that the proportion of males and females with unstranded DNA is not significantly different

***


```{r nodes and metastasis}
#counting number of samples without cancer in nodes
MaleNodes<- nrow(Male[Male$N=="0",])
FemaleNodes<- nrow(Female[Female$N=="0",])

#counting number of samples without metastasis
MaleMetastasis<- nrow(Male[Male$M=="0",])
FemaleMetastasis<- nrow(Female[Female$M=="0",])
```
## **T tests various cell types**