---
title: "MedicalCovariateAnalysis"
author: "Annika Jorgensen"
date: "2023-10-13"
output: html_document
---
PURPOSE: To run statistical analyses on all metadata traits, to see if they are significant between male and females.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Libraries**
The first chunk of code is dedicated to installing the libraries. These libraries are to help execute the differential analysis and helps visualize the data. The code was not included for concision.  
```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)
```

**Read in data**

This code is where you read in all the data files that are going to be used in the script. The data is also converted into a variety of variables that makes the data easier to handle. The data is also cleaned up to make sure the analysis done later is accurate and precise. 
```{r Reading_in_data}
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",") #changing the name of the file
tumorAdjacentExp <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE) #changing the name of the file 
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp)) #changing the column names 

# Importing gene annotations
#genes <- read.table("gencode.v25.chr_patch_hapl_scaff.annotation.bed", header=FALSE, sep="\t")
genes <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)

# Removing Samples due to low quality
metadata <- metadata[!(metadata$ID == "RK023") , ]
metadata <- metadata[!(metadata$ID == "RK106") , ]
metadata <- metadata[!(metadata$ID == "RK113") , ]
metadata <- metadata[!(metadata$ID == "RK135") , ]
metadata <- metadata[!(metadata$ID == "RK105") , ]
metadata <- metadata[!(metadata$ID == "RK116") , ]
metadata <- metadata[!(metadata$ID == "RK066") , ]
metadata <- metadata[!(metadata$ID == "RK096") , ]

#Removing both and NBNC samples
metadata <- metadata[!(metadata$Virus_infection == "NBNC"), ]
metadata <- metadata[!(metadata$Virus_infection == "both"), ]

# Subsetting and ordering metadata to match the count matrix
tumorAdjacentExpSubset <- tumorAdjacentExp[,colnames(tumorAdjacentExp) %in% metadata$sampleid]
metadataSubset <- metadata[metadata$sampleid %in% colnames(tumorAdjacentExpSubset),]
metadataSubset <- metadataSubset[match(colnames(tumorAdjacentExpSubset), metadataSubset$sampleid),]
rownames(metadataSubset) <- metadataSubset$sampleid

# Adding tissue type, converting categorical variables to factors
metadataSubset$tumor <- as.numeric(grepl('tumor', metadataSubset$sampleid, ignore.case=T))

#Swapping lesion type for sample RK169
metadataSubset["RK169-tumor-XY","tumor"] <- 0
metadataSubset["RK169-adjacent-XY","tumor"] <- 1

#Changing rownames to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

#Swapping lesion type for sample RK065
metadataSubset["RK065-tumor-XX","tumor"] <- 0
metadataSubset["RK065-adjacent-XX","tumor"] <- 1

#Changing rownames in metadata to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

metadataSubset$gender_tissue <- paste(metadataSubset$Gender, metadataSubset$tumor, sep="_")
metadataSubset$gender_tissue_viral <- paste(metadataSubset$gender_tissue, metadataSubset$Virus_infection, sep="_")
metadataSubset$library_type <- metadataSubset$strandedness
metadataSubset$library_type <- factor(metadataSubset$library_type)
metadataSubset$tumor <- factor(metadataSubset$tumor)
metadataSubset$Ta <- factor(metadataSubset$Ta)
metadataSubset$Portal_vein_invasion <- factor(metadataSubset$Portal_vein_invasion)
metadataSubset$Hepatic_vein_invasion <- factor(metadataSubset$Hepatic_vein_invasion)
metadataSubset$Bile_duct_invasion <- factor(metadataSubset$Bile_duct_invasion)
metadataSubset$Liver_fibrosisc <- factor(metadataSubset$Liver_fibrosisc)
metadataSubset$Prognosis <- factor(metadataSubset$Prognosis)

write.csv(x=metadataSubset,file="~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/MedicalCovariateAnalysis/data/metadataSubset.csv",row.names=FALSE)
```

**Subset data for statistical analysis**

This code chunk for subsetting data  by male and female tumor tumor adjacent in order to conduct t-tests
```{r subset for stat analy}
#subset metadataSubset data frame by male samples
Male <- metadataSubset[grepl("XY",rownames(metadataSubset)),]

#subset metadataSubset data frame by male tumor samples
MaleTumor<-  Male[grepl("tumor",rownames(Male)),]

#subset metadataSubset by male adjacent samples
MaleAdjacent <- Male[grepl("adjacent", rownames(Male)),]

#subset metadataSubset data frame by female samples
Female <- metadataSubset[grepl("XX",rownames(metadataSubset)),]

#subset metadataSubset data frame female tumor samples
FemaleTumor <- Female[grepl("tumor",rownames(Female)),]

#subset metadataSubset data frame female tumor adjacent samples
FemaleAdjacent <- Female[grepl("adjacent", rownames(Female)),]

```

**F and T test age of samples**
Here is a f test conducted on the ages of the sample
```{r f test age}
#subsetting age of male samples
MaleAge<- Male$Age

#subsetting age of female samples
FemaleAge<- Female$Age

#f test alpha=0.05
var.test(MaleAge, FemaleAge) #Fail to reject null no evidence variance different

#variances equal 
```

No evidence that variances are equal. So, we will conduct a t test with equal variance

T test conducted on ages 
```{r t test age }
#t test alpha=0.05
t.test(MaleAge,FemaleAge,var.equal=TRUE) #Reject null age of male significantly different than females

#means significantly different
```

The ages of male samples compared to female samples is significantly different. It should be noted that the mean age of males is 65.23737 and the mean age of females is 68.76250 which is relatively close. 

CONCLUSION: With 95% confidence we can say that the mean age of male patients compared to female patients is significantly different. 

**F and T test tumor size in mm**
```{r f test tumor size}
#subsetting by tumor size in female tumor samples 
FemaleSize <- FemaleTumor$Tumor_size_mm

#subsetting by tumor size in male tumor samples
MaleSize<- MaleTumor$Tumor_size_mm

#f test alpha=0.05
var.test(MaleSize, FemaleSize) #reject null variances significantly different 

#variances unequal
```
Variances in male tumor size in mm is significantly different than variance in female tumor size. So, we will conduce a t test with unequal variance.
```{r t test tumor size}
#t test tumor size alpha=0.05
t.test(MaleSize,FemaleSize) #Do not reject null no evidence tumor size is different

#no significant difference
```

The tumor size is not significantly different between males and females. 

CONCLUSION: With 95% confidence we conclude that the mean tumor size (in mm) is not significantly different between males and females.

**F and T test overall survival month**
```{r f test overall survival month}
#subsetting overall survival month from male tumor sample
MaleSurvivalMonth<- MaleTumor$Overall_survival_month

#subsetting overall survival month from female tumor sample
FemaleSurvivalMonth<- FemaleTumor$Overall_survival_month

#f test overall survival month alpha=0.05
var.test(MaleSurvivalMonth, FemaleSurvivalMonth) #fail to reject null no evidence variances are different

#variances are equal
```

There is no evidence that the variances are equal, therefore we will conduct a t test where the variances are equal.

```{r t test overall survival month}
#t test overall survival month alpha=0.05
t.test(MaleSurvivalMonth, FemaleSurvivalMonth, var.equal = TRUE) #Reject null survival month significantly different

#significant difference
```

Significant difference between the mean survival month of males compared to females

CONCLUSION: With 95% confidence we conclude that the mean survival month of males is significantly different than the mean survival month of females. 

Also, let it be known that the mean survival month of females is five month shorter than the mean survival month of males. This is inconsistent with the literature.

**CHI-square Stage**
Creating a data frame of counts of males and females with each stage
```{r counts}

```