---
title: "Supplementary Note 2"
author: "Annika Jorgensen"
date: "2023-12-11"
output: html_document
---
PURPOSE: This document is to conduct statistical testing to see if any variables in the metadata are significantly different between males and females. For variables deemed significant further investigation will be done via MDS analysis and these variables may be incorportated as covariates in our DEA. 

METHODS: Using R functions, statistical testing will be done on variables found in the metadata. The following variables will be testing 

-- Age
-- Tumor Stage
-- Nodes
-- Metadata
-- Edmondson Grade
-- Tumor stage in mm
-- Liver Fibrosis
-- Alcohol Intake
-- Smoking
-- Prognosis
-- Survival Months

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(ggplot2)
```

Here the full metadata is read in as a dataframe titled "metadata" and a curated metadata set is titled "metadatasmall"
```{r  Reading in data}
#reading in Full meta dataset
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/Full_RIKEN_clinical_data.csv", row.names=1,header=TRUE, sep=",")

#reading in meta dataset
metadataSmall <- read.csv("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/MedicalCovariateAnalysis/data/metadataSubset.csv", row.names=1)

```


Subsetting metadata such that analyses can be conducted. The curated metadata set only has the samples in the analysis that we want to analyze. However, the curated metadata doesn't contain the full metadata. So, the full metadata set was subsetted for only the samples that we want to study.
```{r subsetting data}
#subsetting full metadata for samples used in analysis
metadataAnalysis <- metadata[rownames(metadata) %in% metadataSmall$ID,]

#subsetting male samples 
Male <- metadataAnalysis[which(metadataAnalysis$Gender== "M"),]

#subsetting female samples 
Female <- metadataAnalysis[which(metadataAnalysis$Gender== "F"),]
```


### **F and T test age of males and females**
Age is a numerical variable so a two-sample t test will be used to identify if the variable is statistically significant. To figure out the type of t test we want to use an f test is used to determine whether or not the variances between the two datasets are equal or not. A significance level of 0.05 is used.
```{r Age difference data}
#subsetting age of male samples
MaleAge<- Male$Age

#subsetting age of female samples
FemaleAge<- Female$Age

#f test alpha=0.05
var.test(MaleAge, FemaleAge) #Fail to reject null no evidence variance different

#variances equal
```

The p value is 0.1851 so therefore we fail to reject the null hypothesis. So, we will conduct a t test with equal variance.

```{r t test age }
#t test alpha=0.05
t.test(MaleAge,FemaleAge,var.equal=TRUE) #Fail to reject null mean age of males and females not significantly different
```

The mean age of males is 65.61 years and the mean age of females is 68.47. The p value is 0.1218 so we fail to reject the null hypothesis.

CONCLUSION: We are 95% confident that the mean age of male patients compared to female patients is not significantly different. This variable will not be further analyzed. 

## **Tumor Stage CHI-square**
Since Tumor Stage is a categorical variable a CHI square homogenity of proportions test was conducted using the chisq.test function.
```{r counts}
#male stage separated by severity 
Male1st<- nrow(Male[Male$Ta== "1", ])
Male2st<- nrow(Male[Male$Ta== "2", ])
Male3st<- nrow(Male[Male$Ta== "3", ])
Male4st<- nrow(Male[Male$Ta== "4", ])

#female stage separated by severity
Female1st<- nrow(Female[Female$Ta== "1", ])
Female2st<- nrow(Female[Female$Ta== "2", ])
Female3st<- nrow(Female[Female$Ta== "3", ])
Female4st<- nrow(Female[Female$Ta== "4", ])

#counts of tumor stage
TumorStage<- data.frame(Males= c(Male1st,Male2st,Male3st,Male4st), Females= c(Female1st, Female2st, Female3st,Female4st))
rownames(TumorStage)<- c(1,2,3,4)

chisq.test(TumorStage)

#using different integrator because of warning
chisq.test(TumorStage, simulate.p.value = TRUE, B = 10000)
```
CONCLUSION: We are 95 percent confident that the proportion of males and females at each Tumor Stage is the same. This variable does not need further analysis.

Here is the table of counts for the patients at each Tumor stage. The row names are the stage.
```{r tumor stage counts}

#sums of rows in TumorStage
TumorStageCounts<-c(27,72,42,9)

#add sums of rows to TumorSTage
TumorStage<- cbind(TumorStage,TumorStageCounts)

#sums of columns in TumorStage
StageTotalCounts<-c(108,42,150)

#add sums of column to TumorStage
TumorStage<- rbind(TumorStage,StageTotalCounts)
rownames(TumorStage)<- c("1","2","3","4","Totals")
```
```{r stage output}
TumorStage
```
Here is a paired percentage bar graph showing the percent of male and females samples in each tumor stage.
```{r percentage graph}

# Sample data
data <- data.frame(
  Sex = rep(c("Males", "Females"), each = 4),
  Category = rep(c("Stage 1", "Stage 2", "Stage 3", "Stage 4"), times = 2),
  Percentage = c(17, 48, 36, 7, 10, 24, 6, 2)
)

custom_colors <- c("#40B0A6", "#FFC20A")

# Calculate the percentages within each group
data <- transform(data, Percentage = Percentage / tapply(Percentage, Sex, sum)[Sex] * 100)

# Create a percentage paired bar graph
ggplot(data, aes(x = Category, y = Percentage, fill = Sex, group = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage Bar Graph Tumor Stage",
       x = "Tumor Stage",
       y = "Percentage") +geom_text(aes(label = sprintf("%.1f%%", Percentage)),
            position = position_dodge(width = 0.9),    # Adjust the width as needed
            vjust = -0.5, hjust = 0.5, size = 3) +  scale_fill_manual(values = custom_colors) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```
## **Investigating nodes and metastasis**
We wanted to see if there were any number of patients with cancer in the lymph nodes or metastasized cancer.
```{r nodes and metastasis}
#counting number of samples without cancer in nodes
MaleNodes<- nrow(Male[Male$N=="1",])
FemaleNodes<- nrow(Female[Female$N=="1",])
print(MaleNodes)
print(FemaleNodes)
#counting number of samples without metastasis
MaleMetastasis<- nrow(Male[Male$M=="1",])
FemaleMetastasis<- nrow(Female[Female$M=="1",])
print(MaleMetastasis)
print(FemaleMetastasis)
```
CONCLUSION: There are no patients with metastastized cancer or cancer in their lymph nodes. This variable does not need further analysis.

## **CHI square Edmondson Grade**
Since Edmondson Grade is a categorical variable a CHI square homogenity of proportions test was conducted using the chisq.test function.
```{r edmondson grade}

#replacing NAs with Zeros. Assuming that NA means that the tumor was not graded
Male[is.na(Male)] <-0
Female[is.na(Female)] <-0

#male grade separated by severity 
MaleEg0<- nrow(Male[Male$Edmondson.grade=="0", ])
MaleEg1<- nrow(Male[Male$Edmondson.grade=="1", ])
MaleEg1.5<- nrow(Male[Male$Edmondson.grade=="1~2", ])
MaleEg2<- nrow(Male[Male$Edmondson.grade=="2", ])
MaleEg2.5<- nrow(Male[Male$Edmondson.grade=="2~3", ])
MaleEg3<- nrow(Male[Male$Edmondson.grade=="3", ])
MaleEg4<- nrow(Male[Male$Edmondson.grade=="4", ])

#female grade separated by severity 
FemaleEg0<- nrow(Female[Female$Edmondson.grade=="0", ])
FemaleEg1<- nrow(Female[Female$Edmondson.grade=="1", ])
FemaleEg1.5<- nrow(Female[Female$Edmondson.grade=="1~2", ])
FemaleEg2<- nrow(Female[Female$Edmondson.grade=="2", ])
FemaleEg2.5<- nrow(Female[Female$Edmondson.grade=="2~3", ])
FemaleEg3<- nrow(Female[Female$Edmondson.grade=="3", ])
FemaleEg4<- nrow(Female[Female$Edmondson.grade=="4", ])

#creating dataframe for edmondson grade
EdmondsonGrade<- data.frame(Males= c(MaleEg0,MaleEg1,MaleEg1.5,MaleEg2,MaleEg2.5,MaleEg3,MaleEg4), Females= c(FemaleEg0,FemaleEg1, FemaleEg1.5, FemaleEg2,FemaleEg2.5,FemaleEg3,FemaleEg4))
rownames(EdmondsonGrade)<- c(0,1,1.5,2,2.5,3,4)

chisq.test(EdmondsonGrade)

#ran using different integrator because of warning
chisq.test(EdmondsonGrade, simulate.p.value = TRUE, B = 10000)
```

CONCLUSION: We are 95 percent confident that the proportion of males and females at each Edmondson Grade is the same. This variable does not need further analysis.

Listed below is the table show the counts of males and females with each Edmondson grade The row names are the levels of Edmondson Grade
```{r counts Edmondson grade}
#sums of rows in EdmondsonGrade
EdmondsonGradeCounts<-c(1,14,8,87,23,16,1)

#add sums of rows to EdmondsonGrade
EdmondsonGrade<- cbind(EdmondsonGrade,EdmondsonGradeCounts)

#sums of columns in EdmondsonGrade
GradeTotalCounts<-c(108,42,150)

#add sums of column to EdmondsonGrade
EdmondsonGrade<- rbind(EdmondsonGrade,GradeTotalCounts)
rownames(EdmondsonGrade)<- c("0","1","1.5","2","2.5","3","4","Totals")
```
```{r grade output}
EdmondsonGrade
```

Below is a percentage paired bar plot showing the percentage of male and female samples at each Edmondson Grade.
```{r percentage graph}

# Sample data
data <- data.frame(
  Sex = rep(c("Males", "Females"), each = 7),
  Category = rep(c("Grade 0", "Grade 1", "Grade 1.5", "Grade 2", "Grade 2.5", "Grade 3", "Grade 4"), times = 2),
  Percentage = c(1, 11, 5, 64, 14, 12, 1, 0, 3, 3, 23, 9, 4, 0)
)

custom_colors <- c("#40B0A6", "#FFC20A")

# Calculate the percentages within each group
data <- transform(data, Percentage = Percentage / tapply(Percentage, Sex, sum)[Sex] * 100)

# Create a percentage paired bar graph
ggplot(data, aes(x = Category, y = Percentage, fill = Sex, group = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage Bar Graph Edmondson Grade",
       x = "Edmondson Grade",
       y = "Percentage") +geom_text(aes(label = sprintf("%.1f%%", Percentage)),
            position = position_dodge(width = 0.9),    # Adjust the width as needed
            vjust = -0.5, hjust = 0.5, size = 3) +  scale_fill_manual(values = custom_colors) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```
### **F and T test tumor size in mm**
Tumor Size is a numerical variable so a two-sample t test will be used to identify if the variable is statistically significant. To figure out the type of t test we want to use an f test is used to determine whether or not the variances between the two datasets are equal or not. A significance level of 0.05 is used.
```{r f test tumor size}
#subsetting by tumor size in female samples 
FemaleSize <- Female$Tumor.size..mm.

#subsetting by tumor size in male samples
MaleSize<- Male$Tumor.size..mm.

#f test alpha=0.05
var.test(MaleSize, FemaleSize) #reject null variances significantly different 

#variances unequal
```

The p value is 9.361e-05 so therefore we reject the null hypothesis. So, we will conduct a t test with unequal variances.

```{r t test tumor size}
#t test tumor size alpha=0.05
t.test(MaleSize,FemaleSize) #Do not reject null no evidence tumor size is different

#no significant difference
```
The mean male tumor size is 41.314 and the mean female tumor size is 33.52. The p value is 0.128 so we fail to reject null hypothesis.

CONCLUSION:We are 95% confident that the mean tumor size (in mm) is not significantly different between males and females. We will not conduct further analysis on this variable.

## **Liver Fibrosis CHI-square**
Since Liver Fibrosis is a categorical variable a CHI square homogenity of proportions test was conducted using the chisq.test function.
```{r Liver Fibrosis}
#male fibrosis separated by severity 
Male0F<- nrow(Male[Male$Liver.fibrosisc== "0", ])
Male1F<- nrow(Male[Male$Liver.fibrosisc== "1", ])
Male2F<- nrow(Male[Male$Liver.fibrosisc== "2", ])
Male3F<- nrow(Male[Male$Liver.fibrosisc== "3", ])
Male4F<- nrow(Male[Male$Liver.fibrosisc== "4", ])

#female fibrosis
Female0F<- nrow(Female[Female$Liver.fibrosisc== "0", ])
Female1F<- nrow(Female[Female$Liver.fibrosisc== "1", ])
Female2F<- nrow(Female[Female$Liver.fibrosisc== "2", ])
Female3F<- nrow(Female[Female$Liver.fibrosisc== "3", ])
Female4F<- nrow(Female[Female$Liver.fibrosisc== "4", ])

#counts of liver fibrosis

LiverFibrosis<- data.frame(Males= c(Male0F,Male1F,Male2F,Male3F,Male4F), Females= c(Female0F, Female1F, Female2F,Female3F,Female4F))

chisq.test(LiverFibrosis)

#using different integrator because of warning
chisq.test(LiverFibrosis, simulate.p.value = TRUE, B = 10000)
```
CONCLUSION: We are 95 percent confident that the proportion of males and females with liver fibrosis is the same. This variable does not need further analysis.

Listed below is the table show the counts of males and females with each level of liver fibrosis. The row names are the severity of liver fibrosis
```{r total counts}
#sums of rows in LiverFibrosis
LiverFibrosisCounts<-c(5,16,30,40,59)

#add sums of rows to LiverFibrosis
LiverFibrosis<- cbind(LiverFibrosis,LiverFibrosisCounts)

#sums of columns in LiverFibrosis
LiverTotalCounts<-c(108,42,150)

#add sums of column to LiverFibrosis
LiverFibrosis<- rbind(LiverFibrosis,LiverTotalCounts)
rownames(LiverFibrosis)<- c("0","1","2","3","4","Totals")
```

```{r librosis counts}
LiverFibrosis
```
Below is a percentage paired bar plot showing the percentage of male and female samples at each level of Liver Fibrosis.
```{r percentage graph}

# Sample data
data <- data.frame(
  Sex = rep(c("Males", "Females"), each = 5),
  Category = rep(c("Severity 0", "Severity 1", "Severity 2", "Severity 3", "Severity 4"), times = 2),
  Percentage = c(5, 13, 25, 26, 39, 0, 3, 5, 14, 20)
)

custom_colors <- c("#40B0A6", "#FFC20A")

# Calculate the percentages within each group
data <- transform(data, Percentage = Percentage / tapply(Percentage, Sex, sum)[Sex] * 100)

# Create a percentage paired bar graph
ggplot(data, aes(x = Category, y = Percentage, fill = Sex, group = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage Bar Graph Liver Fibrosis",
       x = "Severity",
       y = "Percentage") +geom_text(aes(label = sprintf("%.1f%%", Percentage)),
            position = position_dodge(width = 0.9),    # Adjust the width as needed
            vjust = -0.5, hjust = 0.5, size = 3) +  scale_fill_manual(values = custom_colors) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```
## **CHI-squared Alcohol**
Since Alcohol Intake is a categorical variable a Chi square test will be conducted
```{r Alcohol}
#male alcohol intake separated by severity 
Male0<- nrow(Male[Male$Alcohol.intakee== "0", ])
Male1<- nrow(Male[Male$Alcohol.intakee== "1", ])
Male2<- nrow(Male[Male$Alcohol.intakee== "2", ])
Male3<- nrow(Male[Male$Alcohol.intakee== "3", ])
MaleAlcohol<- as.data.frame(cbind(Male0,Male1,Male2,Male3))

#female alcohol
Female0<- nrow(Female[Female$Alcohol.intakee== "0", ])
Female1<- nrow(Female[Female$Alcohol.intakee== "1", ])
Female2<- nrow(Female[Female$Alcohol.intakee== "2", ])
Female3<- nrow(Female[Female$Alcohol.intakee== "3", ])
FemaleAlcohol<- as.data.frame(cbind(Female0,Female1,Female2,Female3))

#counts of alcohol intake

AlcoholIntake<- as.data.frame(cbind(Males= c(Male0,Male1,Male2,Male3), Females= c(Female0, Female1, Female2,Female3)))
rownames(AlcoholIntake)<- c(0,1,2,3)

chisq.test(AlcoholIntake)
# reject null hypothesis
```

CONCLUSION: We are 95% confident that the proportion of males and females who intake alcohol is different. This variable needs further analysis.

Here are two histograms showing the raw counts of males and females at each level of alcohol intake.  
```{r Alcohol Intake Histograme}
#Alcohol Intake Male Histogram
barplot(AlcoholIntake$Males,ylim = c(0,40),
        names.arg = c(0, 1, 2, 3),  # X-axis labels
        col = "skyblue",            # Bar color
        border = "black",           # Bar border color
        main = "Male Alcohol Intake Histogram", # title name
        xlab = "Alcohol Intake Severity",
        ylab = "Counts",
        space = 0 ) 

# Female Alcohol Intake  Histogram

barplot(AlcoholIntake$Females, ylim= c(0,40),
        names.arg = c(0, 1, 2, 3),  # X-axis labels
        col = "skyblue",            # Bar color
        border = "black",           # Bar border color
        main = "Female Alcohol Intake Histogram", # title name
        xlab = "Alcohol Intake Severity",
        ylab = "Counts",
        space = 0 )   
```
Here is a sample table for males and female patients alcohol intake. The level of severity is the row name.
```{r total columns alcohol}
#sums of rows in Alcoholintake
AlcoholTotalCounts<-c(63,25,42,20)

#add sums of rows to AlcoholIntake
AlcoholIntake<- cbind(AlcoholIntake,AlcoholTotalCounts)

#sums of columns in AlcoholIntake
AlcoholTotalCounts<-c(108,42,150)

#add sums of column to AlcoholIntake
AlcoholIntake<- rbind(AlcoholIntake,AlcoholTotalCounts)
rownames(AlcoholIntake)<- c("0","1","2","3","Totals")
```

```{r output alcohol intake}
AlcoholIntake
```
Below is a percentage paired bar plot showing the percentage of male and female sample at each level of alcohol intake.
```{r percentage graph}

# Sample data
data <- data.frame(
  Sex = rep(c("Males", "Females"), each = 4),
  Category = rep(c("Level 0", "Level 1", "Level 2", "Level 3"), times = 2),
  Percentage = c(31, 20, 37, 20, 32, 5, 5, 0)
)

custom_colors <- c("#40B0A6", "#FFC20A")

# Calculate the percentages within each group
data <- transform(data, Percentage = Percentage / tapply(Percentage, Sex, sum)[Sex] * 100)

# Create a percentage paired bar graph
ggplot(data, aes(x = Category, y = Percentage, fill = Sex, group = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage Bar Graph Alcohol Intake",
       x = "Level of Intake",
       y = "Percentage") +geom_text(aes(label = sprintf("%.1f%%", Percentage)),
            position = position_dodge(width = 0.9),    # Adjust the width as needed
            vjust = -0.5, hjust = 0.5, size = 3) +  scale_fill_manual(values = custom_colors) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```
### **Test of Proportions Smoking Males and Females** 
Smoking is a dichotomous variable so we will conduct a test of proportions
```{r smoking z test}
#proportion of male smokers
MaleSmokers<-sum(Male$Smoking)
MaleSmokers


#proportion of female smokers
FemaleSmokers<- sum(Female$Smoking)
FemaleSmokers

```

```{r two sample z test}
# alpha=0.05
prop.test(x=c(MaleSmokers,FemaleSmokers), n= c(108,42), p = NULL, alternative = "two.sided", correct = TRUE) #reject null hypothesis that there is no difference between the number of men and women that smoke

#the number of men and women that smoke is significantly different
```

CONCLUSION: We are 95% confident that the proportion of males and females who smoke are different. Since 77% of males smoke and 9% of females smoke it is clear that more males that females smoke.

## **CHI square Prognosis**
Sincle prognosis is a categorical variable a chi square homogenity of proportions will be conducted.
```{r Prognosis}

#male prognosis separated by severity?
MaleProg0<- nrow(Male[Male$Prognosisf=="0",])
MaleProg1<- nrow(Male[Male$Prognosisf=="1",])
MaleProg2<- nrow(Male[Male$Prognosisf=="2",])
MaleProg3<- nrow(Male[Male$Prognosisf=="3",])

#female prognosis separated by severity?
FemaleProg0<- nrow(Female[Female$Prognosisf=="0",])
FemaleProg1<- nrow(Female[Female$Prognosisf=="1",])
FemaleProg2<- nrow(Female[Female$Prognosisf=="2",])
FemaleProg3<- nrow(Female[Female$Prognosisf=="3",])

#counts of Prognosis
Prognosis<- as.data.frame(cbind(Males= c(MaleProg0,MaleProg1,MaleProg2,MaleProg3), Females= c(FemaleProg0, FemaleProg1,FemaleProg2,FemaleProg3)))

chisq.test(Prognosis)

#using different integrator because of warning
chisq.test(Prognosis, simulate.p.value = TRUE, B = 10000)
```

CONCLUSION: We are 95% confident that the proportion of males and females at each level of prognosis is the same

Here is a sample table with number of patients at each level of prognosis. The level of severity is the row name.
```{r prognosis sample table}
#sums of rows in Prognosis
PrognosisCounts<-c(124,21,3,2)

#add sums of rows to Prognosis
Prognosis<- cbind(Prognosis,PrognosisCounts)

#sums of columns in Prognosis
PrognosisTotalCounts<-c(108,42,150)

#add sums of column to Prognosis
Prognosis<- rbind(Prognosis,PrognosisTotalCounts)
rownames(Prognosis)<- c("0","1","2","3","Totals")
```

```{r prognosis output}
Prognosis
```
```{r percentage graph}

# Sample data
data <- data.frame(
  Sex = rep(c("Males", "Females"), each = 4),
  Category = rep(c("Level 0", "Level 1", "Level 2", "Level 3"), times = 2),
  Percentage = c(92, 13, 2, 1, 32, 8, 1, 1)
)

custom_colors <- c("#40B0A6", "#FFC20A")

# Calculate the percentages within each group
data <- transform(data, Percentage = Percentage / tapply(Percentage, Sex, sum)[Sex] * 100)

# Create a percentage paired bar graph
ggplot(data, aes(x = Category, y = Percentage, fill = Sex, group = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage Bar Graph Prognosis",
       x = "Level of Severity",
       y = "Percentage") +geom_text(aes(label = sprintf("%.1f%%", Percentage)),
            position = position_dodge(width = 0.9),    # Adjust the width as needed
            vjust = -0.5, hjust = 0.5, size = 3) +  scale_fill_manual(values = custom_colors) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```

### **F and T test overall survival month**
Since survival month is a numerical variable a two sample t test will be conducted. To see what type of t test shall be used a f test is conducted first
```{r f test overall survival month}
#subsetting overall survival month from male sample
MaleSurvivalMonth<- Male$Overall.survival..month.

#subsetting overall survival month from female sample
FemaleSurvivalMonth<- Female$Overall.survival..month.

#f test overall survival month alpha=0.05
var.test(MaleSurvivalMonth, FemaleSurvivalMonth) #fail to reject null no evidence variances are different

#variances are equal
```

The p value is 0.3362 so we fail to reject the null hypothesis. A two sample t test of equal variance will be conducted.

```{r t test overall survival month}
#t test overall survival month alpha=0.05
t.test(MaleSurvivalMonth, FemaleSurvivalMonth, var.equal = TRUE) #Reject null survival month significantly different

#significant difference
```
The mean overall survival month of males is 29.2 months and the mean overall survival month of females is 23.9 months. Note that the survival month for females is shorter than males which is atypical. 

CONCLUSION: We are 95% confident that the mean survival month of males is different that the mean survival month of females. 