---
title: "Viral Etiology Sex Covariate"
author: "Annika Jorgensen"
date: "2023-01-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **R Markdown**

Title: "DEA Viral Etiology with Sex as Covariate"  
Author: Annika Jorgensen  
Date: 01/04/2023 
Purpose: The Purpose of this file is to do a DEA with the ICGC liver cancer data stratified by etiology with sex as  a covariate. The purpose is to make sure that sex is account for sex as a potential influence of the expressed genes. 

### **Libraries**

The first chunk of code is dedicated to installing the libraries. These libraries are to help execute the differential analysis and helps visualize the data. The code was not included for concision.   
```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)
```

**Defining Colors** 

This chunk defines color palette variables that are going to be used in plots later on the script. These variables are defined by converting BrewerCode palettes into palettes that can be used in R. 
```{r Color_Palettes}
viralPalette <- brewer.pal(8, "Set1")
hbvColor <- viralPalette[1]
hcvColor <- viralPalette[2]
bothColor <- viralPalette[3]
neitherColor <- viralPalette[4]

sexTissuePalette <- brewer.pal(12, "Paired")
maleTumorColor <- sexTissuePalette[4]
maleAdjacentColor <- sexTissuePalette[3]
femaleTumorColor <- sexTissuePalette[6]
femaleAdjacentColor <- sexTissuePalette[5]
```

**Read in data**

This code is where you read in all the data files that are going to be used in the script. The data is also converted into a variety of variables that makes the data easier to handle. The data is also cleaned up to make sure the analysis done later is accurate and precise. 
```{r Reading_in_data}
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",") #changing the name of the file
tumorAdjacentExp <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE) #changing the name of the file 
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp)) #changing the column names 

# Importing gene annotations
#genes <- read.table("gencode.v25.chr_patch_hapl_scaff.annotation.bed", header=FALSE, sep="\t")
genes <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)

# Removing Samples due to low quality
metadata <- metadata[!(metadata$ID == "RK023") , ]
metadata <- metadata[!(metadata$ID == "RK106") , ]
metadata <- metadata[!(metadata$ID == "RK113") , ]
metadata <- metadata[!(metadata$ID == "RK135") , ]
metadata <- metadata[!(metadata$ID == "RK105") , ]
metadata <- metadata[!(metadata$ID == "RK116") , ]
metadata <- metadata[!(metadata$ID == "RK066") , ]
metadata <- metadata[!(metadata$ID == "RK096") , ]

#Removing both and NBNC samples
metadata <- metadata[!(metadata$Virus_infection == "NBNC"), ]
metadata <- metadata[!(metadata$Virus_infection == "both"), ]

# Subsetting and ordering metadata to match the count matrix
tumorAdjacentExpSubset <- tumorAdjacentExp[,colnames(tumorAdjacentExp) %in% metadata$sampleid]
metadataSubset <- metadata[metadata$sampleid %in% colnames(tumorAdjacentExpSubset),]
metadataSubset <- metadataSubset[match(colnames(tumorAdjacentExpSubset), metadataSubset$sampleid),]
rownames(metadataSubset) <- metadataSubset$sampleid

# Adding tissue type, converting categorical variables to factors
metadataSubset$tumor <- as.numeric(grepl('tumor', metadataSubset$sampleid, ignore.case=T))

#Swapping lesion type for sample RK169
metadataSubset["RK169-tumor-XY","tumor"] <- 0
metadataSubset["RK169-adjacent-XY","tumor"] <- 1

#Changing rownames to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

#Swapping lesion type for sample RK065
metadataSubset["RK065-tumor-XX","tumor"] <- 0
metadataSubset["RK065-adjacent-XX","tumor"] <- 1

#Changing rownames in metadata to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

metadataSubset$gender_tissue <- paste(metadataSubset$Gender, metadataSubset$tumor, sep="_")
metadataSubset$gender_tissue_viral <- paste(metadataSubset$gender_tissue, metadataSubset$Virus_infection, sep="_")
metadataSubset$library_type <- metadataSubset$strandedness
metadataSubset$library_type <- factor(metadataSubset$library_type)
metadataSubset$tumor <- factor(metadataSubset$tumor)
metadataSubset$Ta <- factor(metadataSubset$Ta)
metadataSubset$Portal_vein_invasion <- factor(metadataSubset$Portal_vein_invasion)
metadataSubset$Hepatic_vein_invasion <- factor(metadataSubset$Hepatic_vein_invasion)
metadataSubset$Bile_duct_invasion <- factor(metadataSubset$Bile_duct_invasion)
metadataSubset$Liver_fibrosisc <- factor(metadataSubset$Liver_fibrosisc)
metadataSubset$Prognosis <- factor(metadataSubset$Prognosis)
```

```{r DGE Object}
# Creating the DGEList object
dge <- DGEList(counts=tumorAdjacentExpSubset, genes=genes)
colnames(dge) <- colnames(tumorAdjacentExpSubset)
dge$samples$sex <- metadataSubset$Gender
dge$samples$viral <- paste(metadataSubset$Virus_infection, metadataSubset$tumor, sep="_")
dge$samples$ID <- metadataSubset$ID
dge$samples$tumor <- metadataSubset$tumor
dge$samples$gender_tissue <- metadataSubset$gender_tissue
dge$samples$gender_tissue_viral <- metadataSubset$gender_tissue_viral
dge$samples$library_type <- metadataSubset$library_type
dge$samples$edmonson_grade <- metadataSubset$Edmondson_grade
dge$samples$Ta <- metadataSubset$Ta
dge$samples$survival <- metadataSubset$Overall_survival_month
dge$samples$smoking <- factor(metadataSubset$Smoking)
dge$samples$alcohol <- factor(metadataSubset$Alcohol_intake)
dge$samples$fibrosis <- factor(metadataSubset$Liver_fibrosisc)

# Inspecting the N of samples in each group
table(dge$samples$gender_tissue_viral)
```

**Calculating fpkm values **

This chunk of code takes the fpkm of all the genes in the dataset and calculates the mean. They also filter out genes that have a fpkm of 0.5 or lower. 

Fpkm stands for  fragments per kilo base of exon per million this term is interchangable with Rpkm (reads per kilobase of exon per million. This measure is a normalization method which allows us to compare gene expression levels by rescaling both library size and gene length.   

Fpkm is calculated by multiplying the number of reads mapped to a gene by 1,000,  1,000,000 and then dividing that number by the total number of mapped reads to gene length in base pairs. 

Please note that that calculation is done for RPKM which is analogous to Fpkm.
```{r fpkm}
# ======================================
# Filtering expression data
# ======================================

# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

## ** **DEA tumor vs. tumor- adjacent **

This section is doing voom/limma DEA on all tumor vs. tumor adjacent samples

```{r First Design Matrix}
# ===========================================================
# ===========================================================
# Analysis of all tumor vs. tumor-adjacent regardless of sex
# ===========================================================
# ===========================================================

# Creating a new design model matrix with the variable of interest and the
# library type
design <- model.matrix(~0+dge$samples$tumor+dge$samples$library_type)

#I just put design matrix in twice because I didn't remember what inputs to put in. 
identical(design,design,num.eq=TRUE)

colnames(design) <- gsub("dge\\$samples\\$tumor", "tumor", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
head(design)
```

**voom**

voom is a function that lies within a package called limma. limma/voom is used in DGE analysis. voom is a function that takes the counts in a metadata set and transforms them into log2 of TMM values calculated above in the normalization factors. A linear model is then fitted to the TMM for each gene and residuals are calculated. A smoothed curve is then fitted to the square root of the residual standard deviation by the average expression (this is the red line). This smooth curve is then used to obtain weights for each gene and sample that are passed into limma along with the TMM values.  
```{r First voom run}
# Running voom again with the new design matrix.
#pdf("~/Desktop/Research Projects/Liver Cancer Project/Figures/etiology tumor tumor adjacent voom.pdf", width=12, height=12)
v <- voomWithQualityWeights(dge, design, plot=TRUE)
#dev.off()
```

**limma**

This sections marks the beginning of running limma. limma creates a linear fit to the data, makes comparisons of the fitted data, and applies Bayes smoothing. limma startes by creating a variable that has all of the duplicate correlation values on v and design. These correlation values will be used later in a linear fit.
```{r Limma fit tumor v. tumor- adjacent}
# ============================================================================================
# Differential expression analysis with limma - all male tumor adjacent vs. non-tumor-adjacent
# ===========================================================================================

# Block design for individual. This is used in tumor-normal comparisons with
# paired samples.
corfit <- duplicateCorrelation(v, design, block = v$targets$ID)
# This should give a positive correlation value. It represents the
# correlation between measurements made on the same person.
corfit$consensus
```

**limma graph**

This is the linear model with limma, notice that the correlation values were the duplicate correlations used earlier
```{r Linear fit limma}
# Fitting the linear model with limma.
# If using paired samples, the within-patient correlation and a block design
# for patient is used to account for pairwise samples
fit <- lmFit(v, design, block = v$targets$ID, correlation = corfit$consensus)
```

**Coefficient vector**

This code chunk involves extracting coefficients from the linear fit model and storing them in a vector for later use.
```{r Design Matrix}
# Contrast design for differential expression
# Defining pairwise comparisons
contrasts <- makeContrasts(Adjacent_vs_Tumor = tumor1 - tumor0,
                           levels=colnames(design))
head(contrasts)
# Assigning all comparisons to a vector for later
allComparisons <- colnames(contrasts)
```

**Contrast Analysis**

This next code chunk reorients the linear model obtained earlier and obtains the coefficients and standard errors from the model. This step also sets us up to apply Empirical Bayes smoothing. 
```{r Contrast Analysis}
# Running contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)
# Look at N of DEGs with adj. p <0.01 and log2FC>2
summary(decideTests(vfit, adjust.method = "BH", p.value = 0.05, lfc = 2))
```

**Bayes smoothing**

This code chunk uses Empirical Bayes smoothing to plot the final model after doing the limma and voom analysis. 

Empirical Bayes smoothing is a way to account for uncertainty. The technique uses the population in a region as a measure of confidence. Meaning that areas with low margin of error are left untouched while estimates with higher margin of error are moved closer to the global average. 

The write.csv is commented out because it is not needed for this file
```{r Bayes Analysis}
# Computing differential expression based on the empirical Bayes moderation of
# the standard errors towards a common value. Robust = should the estimation of
# the empirical Bayes prior parameters be robustified against outlier sample
# variances?
veBayesFit <- eBayes(vfit, robust=TRUE)

plotSA(veBayesFit, main = "Final model: Mean-variance trend")


vTopTable <- topTable(veBayesFit, n=Inf, p.value=1, lfc=0)
DEGs <- topTable(veBayesFit, n=Inf, p.value=0.05, lfc=2)
#DEGs_print <- data.frame(DEGs$GENEID, DEGs$gene_name, DEGs$adj.P.Val, DEGs$logFC)
#write.csv(DEGs_print, "~/R/gene_list_tumor_vs_tumor_adjacent.csv")
```

## **DEA tumor vs. tumor-adjacent etioloy-specific** 
This section is doing DEA on tumor tumor-adj. data that is stratified by etiology.
**Design matrix**

This code chunk is setting up a second design matrix for a voom analysis of tumor vs. non-tumor differentiated by etiology. Please note that sex was added as a covariate.

```{r Second Design Matrix }
# ===========================================================
# ===========================================================
# Analysis of tumor vs. non-tumor differentiated by sex
# ===========================================================
# ===========================================================

# ======================================
# voom transformation
# ======================================

# Creating a design model matrix with the variable of interest

#Added library type and Targets to the design matrix 
#design <- model.matrix(~0+dge$samples$gender_tissue_viral)

design <- model.matrix(~0 + v$targets$viral+v$targets$library_type + v$targets$sex)

identical(design,design,num.eq=TRUE)

#Added target and library type column names
colnames(design) <- gsub("v\\$targets\\$viral", "", colnames(design))
colnames(design) <- gsub("v\\$targets\\$library_typeunstranded", "library_type", colnames(design))
colnames(design) <- gsub("v\\$targets\\$sex", "Sex", colnames(design))
tail(design)
```

***Voom for Male Tumor vs. Tumor adjacent***

This code chunk is using the voom function and outputting the graph. 
```{r Second Voom}
# Running voom with quality weights. Normalizes expression intensities so that
# the log-ratios have similar distributi                     bnons across a set of samples.
# To quantile normalize, add normalize.method="quantile"
# Running parallel
#pdf("~/Desktop/ResearchProjects/LiverCancer/Figures/ etiology tumor tumor adjacent voom.pdf", width=12, height=12)
v <- voomWithQualityWeights(dge, design, plot=TRUE)
#dev.off()
```

***limma***

This code chunk is for doing paired samples correlation tests. The value of the correlation should be positive.
```{r Coorelation Values}
# Block design for individual. This is used in tumor-normal comparisons with
# paired samples.
corfit <- duplicateCorrelation(v, design, block = v$targets$ID)
# This should give a positive correlation value. It represents the
# correlation between measurements made on the same person.
corfit$consensus
```

***limma graph***
This code chuck is doing a linear fit model with limma for male tumor vs tumor-adjacent. 
```{r Linear Fit}

# Fitting the linear model with limma.
# If using paired samples, the within-patient correlation and a block design
# for patient is used to account for pairwise samples
fit <- lmFit(v, design, block = v$targets$ID, correlation = corfit$consensus)
```

***Coefficient vector***

This code chunk does pairwise comparisons of the male tumor vs. tumor adjacent sample for contrast design for differential expressions and stores the comparisons into a vector for later. 
```{r Constrasts again}
# Contrast design for differential expression
# Defining pairwise comparisons

contrasts <- makeContrasts(Adjacent_HBV_vs_Tumor_HBV= HBV_1 - HBV_0,         
                           Adjacent_HCV_vs_Tumor_HCV = HCV_1 - HCV_0,       
                           levels=colnames(design))

#Couldn't make the pair wise comparisions would so I just picked some random column names.
#contrasts <- makeContrasts(Adjacent_F_vs_Tumor_F = F_1_HBV - F_0_HBV,
                           #Adjacement_M_vs_Tumor_M=M_1_both- M_0_both,
                           #levels=colnames(design))


head(contrasts)


# Assigning all comparisons to a vector for later
allComparisons <- colnames(contrasts)
```

```{r Contrast Analysis 2}

# Running contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)
# Looking at N of DEGs with adj. p <0.01 and log2FC>2
summary(decideTests(vfit, adjust.method = "BH", p.value = 0.05, lfc = 2))
```

```{r Bayes fit}
# Computing differential expression based on the empirical Bayes moderation of
# the standard errors towards a common value. Robust = should the estimation of
# the empirical Bayes prior parameters be robustified against outlier sample
# variances?
veBayesFit <- eBayes(vfit, robust=TRUE)

#pdf("~/Desktop/ResearchProjects/LiverCancer Project/Figures/etiology tumor tumor adjacent final model.pdf", width=12, height=12)
plotSA(veBayesFit, main = "Final model: Mean-variance trend")
#dev.off()

DEGs_HBV <- topTable(veBayesFit, coef=1, n=Inf, p.value=0.05, lfc=2)
write.csv(DEGs_HBV, "~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/ViralEtiology/GeneLists/DEG_HBV_covariate.csv")

DEGs_HCV <- topTable(veBayesFit, coef=2, n=Inf, p.value=0.05, lfc=2)
write.csv(DEGs_HCV, "~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/ViralEtiology/GeneLists/DEG_HCV_covariate.csv")
```

```{r shared and unique genes}
#library(arsenal)
#comparedf(x=DEGs_HBV, y=DEGs_HCV, by="gene_name")
#summary(comparedf(x=DEGs_HBV, y=DEGs_HCV))
```

```{r new volcano plot HBV}
vTopTable_HBV <- topTable(veBayesFit, coef=1, n=Inf, p.value=1, lfc=0)

df <- data.frame(vTopTable_HBV$adj.P.Val, vTopTable_HBV$logFC, vTopTable_HBV$chr, vTopTable_HBV$GENEID, vTopTable_HBV$gene_name)
colnames(df) <- c("adj.P.Val", "logFC", "chr", "id", "name")
dfSig <- df[(abs(df$logFC) >= 2 & df$adj.P.Val <= 0.05),]$id

p<-ggplot(data=df, aes(x=logFC, y=-log10(adj.P.Val))) + geom_point(alpha=0.5)+theme_bw()

p2 <-p+ geom_vline(xintercept=c(-2, 2), linetype= "dashed", size=1.5) + geom_hline(yintercept= 2, linetype= "dashed", size=1.5) +  xlim(c(-10, 10)) #+ ylim(c(0, 50))

p3<-p2+ theme(legend.position = c(0.925,0.75)) + geom_point(aes(color=abs(logFC)>=2 & adj.P.Val<=0.05), size=4)+theme(legend.position="none")+scale_color_manual(values = c("black","maroon"))

p4<- p3+theme(axis.title = element_text(size=35), axis.text = element_text(size=35, color="black"), legend.title = element_text(size=35)) + labs(x= "log fold change", y="-log(FDR- adjusted p value)")

p4

pdf( "~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/ViralEtiology/VolcanoPlots/Volcano_plot_HBV.pdf", width=12, height=12)
p4
dev.off()
```

```{r HCV new volcanoplot}
vTopTable_HCV <- topTable(veBayesFit, coef=2, n=Inf, p.value=1, lfc=0)

df <- data.frame(vTopTable_HCV$adj.P.Val, vTopTable_HCV$logFC, vTopTable_HCV$chr, vTopTable_HCV$GENEID, vTopTable_HCV$gene_name)
colnames(df) <- c("adj.P.Val", "logFC", "chr", "id", "name")
dfSig <- df[(abs(df$logFC) >= 2 & df$adj.P.Val <= 0.05),]$id

p<-ggplot(data=df, aes(x=logFC, y=-log10(adj.P.Val))) + geom_point(alpha=0.5)+theme_bw()

p2 <-p+ geom_vline(xintercept=c(-2, 2), linetype= "dashed", size=1.5) + geom_hline(yintercept= 2, linetype= "dashed", size=1.5) +  xlim(c(-10, 10)) #+ ylim(c(0, 50))

p3<-p2+ theme(legend.position = c(0.925,0.75)) + geom_point(aes(color=abs(logFC)>=2 & adj.P.Val<=0.05), size=4)+theme(legend.position="none")+scale_color_manual(values = c("black","maroon"))

p4<- p3+theme(axis.title = element_text(size=35), axis.text = element_text(size=35, color="black"), legend.title = element_text(size=35)) + labs(x= "log fold change", y="-log(FDR- adjusted p value)")

p4

pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/ViralEtiology/VolcanoPlots/Volcano_plot_HCV.pdf", width=12, height=12)
p4
dev.off()
```