---
title: "LogvFoldvChangevPlots"
author: "Annika Jorgensen"
date: "2024-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Libraries, include= FALSE}
#==================================
# Markdown Libraries 
#==================================

#==================================
# DEG Analysis Libraries 
#==================================

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)                                  
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(UpSetR)
library(formatR)
library(tinytex)
```

##**Plotting Original LogFC Plot**
Importing the unfiltered gene lists. 
```{r}
#importing unfiltered gene lists
FemaleUnfilteredGeneList <- read.csv("FemaleTumorTumorAdjacent/Figures/GeneLists/unfiltered_DEGs_F.csv")
MaleUnfilteredGeneList <- read.csv("FemaleTumorTumorAdjacent/Figures/GeneLists/unfiltered_DEGs_M.csv")
OverallUnfilteredGeneList <- read.csv("TumorTumorAdjacent/Figures/GeneLists/unfiltered_gene_list_tumor_vs_tumor_adjacent.csv")
```

Matching gene lists to ensure all genes plotted are consistent across males and females.
```{r}
#matching the male gene list to the female gene list 
FemaleUnfilteredGeneList <- FemaleUnfilteredGeneList[match(MaleUnfilteredGeneList$gene_name, FemaleUnfilteredGeneList$gene_name), ]

#checking if they are identical
identical(FemaleUnfilteredGeneList$gene_name, MaleUnfilteredGeneList$gene_name)
```

```{r}
#plotting logFC should see linear trend
plot(MaleUnfilteredGeneList$logFC, FemaleUnfilteredGeneList$logFC)

# Calculate and display the regression line
regression <- lm(MaleUnfilteredGeneList$logFC~FemaleUnfilteredGeneList$logFC)
abline(regression)
```

##**Plotting LogFC Male and Female Colored by Significance**

```{r}
#creating regression line
regression <- lm(MaleUnfilteredGeneList$logFC~FemaleUnfilteredGeneList$logFC)

#creating data frame for logFC plot.
df <- data.frame(MaleUnfilteredGeneList$logFC, MaleUnfilteredGeneList$gene_name, MaleUnfilteredGeneList$adj.P.Val, FemaleUnfilteredGeneList$logFC, FemaleUnfilteredGeneList$gene_name, FemaleUnfilteredGeneList$adj.P.Val)

#Placing residuals into data frame for logFC plot
df$residuals <- residuals(regression)

#naming columns 
colnames(df) <- c("M_logFC", "M_name", "M_adj.P.Val", "F_logFC", "F_name", "F_adj.P.Val","residuals")
```

```{r}
#plotting logFC and coloring by significance 
plot(df$M_logFC, df$F_logFC, 
     col=ifelse(df$M_adj.P.Val>=0.05 & df$F_adj.P.Val>=0.05, "black", #not sig
                ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val>0.05,"blue", #sig male not sig female
                       ifelse(df$M_adj.P.Val>0.05 & df$F_adj.P.Val<0.05, "maroon", #sig female not sig male
                              ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val<0.05, "goldenrod" , "green"))))) # gold = sig both 
legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Not significant", "Signficiant males only", "Significant Females only", "Signfiicant Males and Females"))
```

```{r}
pdf("LogFCPlotGeneLists/logFcMaleFemaleColoredBySignificanceV3.pdf", width=12, height=12)
#plotting logFC and coloring by significance 
plot(df$M_logFC, df$F_logFC, 
     col=ifelse(df$M_adj.P.Val>=0.05 & df$F_adj.P.Val>=0.05, "black", #not sig
                ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val>0.05,"blue", #sig male not sig female
                       ifelse(df$M_adj.P.Val>0.05 & df$F_adj.P.Val<0.05, "maroon", #sig female not sig male
                              ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val<0.05, "goldenrod" , "green"))))) # gold = sig both 

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)

legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Not significant", "Signficiant males only", "Significant Females only", "Signfiicant Males and Females"))
dev.off()
```

**Coloring by Restricted Significance**

```{r}
# restricting data to only significant genes 
df<-df[which(df$M_adj.P.Val<0.05 | df$F_adj.P.Val < 0.05), ]
```

```{r}

plot(df$M_logFC, df$F_logFC, 
     col=ifelse(df$M_adj.P.Val>=0.05 & df$F_adj.P.Val>=0.05, "black", # not sig (should be nothing)
                ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val>0.05,"blue", # sig male not female
                       ifelse(df$M_adj.P.Val>0.05 & df$F_adj.P.Val<0.05, "maroon", # sig female not male
                              ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val<0.05, "goldenrod", "green")))))  #gold sig both

#adding and vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)

legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Not significant", "Signficiant males only", "Significant Females only", "Signficant Males and Females"))
```
```{r}
pdf("LogFCPlotGeneLists/logFcMaleFemaleColoredByRestrictedSignificanceV3.pdf", width=12, height=12)
plot(df$M_logFC, df$F_logFC, 
     col=ifelse(df$M_adj.P.Val>=0.05 & df$F_adj.P.Val>=0.05, "black", # not sig (should be nothing)
                ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val>0.05,"blue", # sig male not female
                       ifelse(df$M_adj.P.Val>0.05 & df$F_adj.P.Val<0.05, "maroon", # sig female not male
                              ifelse(df$M_adj.P.Val<0.05 & df$F_adj.P.Val<0.05, "goldenrod", "green"))))) #gold sig both

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)

legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Not significant", "Signficiant males only", "Significant Females only", "Signficant Males and Females"))
dev.off()
```

##**Plotting LogFC Male and Female Colored by LogFC values**
```{r}
#creating data frame for logFC plot.
df <- data.frame(MaleUnfilteredGeneList$logFC, MaleUnfilteredGeneList$gene_name, MaleUnfilteredGeneList$adj.P.Val, FemaleUnfilteredGeneList$logFC, FemaleUnfilteredGeneList$gene_name, FemaleUnfilteredGeneList$adj.P.Val)

#Placing residuals into data frame for logFC plot
df$residuals <- residuals(regression)

#naming columns 
colnames(df) <- c("M_logFC", "M_name", "M_adj.P.Val", "F_logFC", "F_name", "F_adj.P.Val","residuals")
```
```{r}
plot(df$M_logFC, df$F_logFC, 
     col = ifelse(df$M_logFC <= 0 & df$F_logFC <= 0,"black", #both logFC < 0
                                    ifelse(df$M_logFC > 0 & df$F_logFC < 0, "blue", #males logFC >0 females logFC < 0
                                      ifelse(df$M_logFC < 0 & df$F_logFC > 0, "maroon", #females logFC > 0 males logFC < 0
                                           ifelse(df$M_logFC > 0 & df$F_logFC > 0, "goldenrod", "green")))))# golden both logFC > 0

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
     
legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Male & Females logFC <= 0", "Males logFC > 0 only", "Females logFC > 0 only", "Males & Females logFC < 0"))
```

```{r}
pdf("LogFCPlotGeneLists/logFcMaleFemaleLogFCValueV3.pdf", width=12, height=12)
plot(df$M_logFC, df$F_logFC, 
     col = ifelse(df$M_logFC <= 0 & df$F_logFC <= 0,"black", #both logFC < 0
                                    ifelse(df$M_logFC > 0 & df$F_logFC < 0, "blue", #males logFC >0 females logFC < 0
                                      ifelse(df$M_logFC < 0 & df$F_logFC > 0, "maroon", #females logFC > 0 males logFC < 0
                                           ifelse(df$M_logFC > 0 & df$F_logFC > 0, "goldenrod", "green")))))# golden both logFC > 0

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
     
legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Male & Females logFC <= 0", "Males logFC > 0 only", "Females logFC > 0 only", "Males & Females logFC < 0"))
dev.off()
```

```{r}
# restricting data to only significant genes 
df<-df[which(df$M_adj.P.Val<0.05 | df$F_adj.P.Val < 0.05), ]
```

```{r}
plot(df$M_logFC, df$F_logFC, 
     col = ifelse(df$M_logFC <= 0 & df$F_logFC <= 0,"black", #both logFC < 0
                                    ifelse(df$M_logFC > 0 & df$F_logFC < 0, "blue", #males logFC >0 females logFC < 0
                                      ifelse(df$M_logFC < 0 & df$F_logFC > 0, "maroon", #females logFC > 0 males logFC < 0
                                           ifelse(df$M_logFC > 0 & df$F_logFC > 0, "goldenrod", "green")))))# golden both logFC > 0

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
     
legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Male & Females logFC <= 0", "Males logFC > 0 only", "Females logFC > 0 only", "Males & Females logFC < 0"))
```

```{r}
pdf("LogFCPlotGeneLists/logFcMaleFemaleRestrictedLogFCValueV3.pdf", width=12, height=12)
plot(df$M_logFC, df$F_logFC, 
     col = ifelse(df$M_logFC <= 0 & df$F_logFC <= 0,"black", #both logFC < 0
                                    ifelse(df$M_logFC > 0 & df$F_logFC < 0, "blue", #males logFC >0 females logFC < 0
                                      ifelse(df$M_logFC < 0 & df$F_logFC > 0, "maroon", #females logFC > 0 males logFC < 0
                                           ifelse(df$M_logFC > 0 & df$F_logFC > 0, "goldenrod", "green")))))# golden both logFC > 0

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
     
legend("topleft", pch=c(1), 
       col=c("black", "blue", "maroon", "goldenrod"),
       legend=c("Male & Females logFC <= 0", "Males logFC > 0 only", "Females logFC > 0 only", "Males & Females logFC < 0"))
dev.off()
```

## **LogFC Plot Genes Unique to Males**

Importing filtered gene lists
```{r}
#importing differentially expressed gene lists LogFC > 2 and p.value < 0.05
FemaleGeneList <- read.csv("FemaleTumorTumorAdjacent/Figures/GeneLists/Female_DEGs.csv")
MaleGeneList <- read.csv("MaleTumorTumorAdjacent/Figures/GeneLists/Male_all_DEGs.csv")
OverallGeneList <- read.csv("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/TumorTumorAdjacent/Figures/GeneLists/gene_list_tumor_vs_tumor_adjacent.csv")
```

Getting gene names unique to males
```{r}
# Extract rows where genes are in male list but not female and overall list 
filtered_rows <- which(!(MaleGeneList$gene_name %in% FemaleGeneList$gene_name | MaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name))

# Extract the p value and logFC based on  of the filtered rows
extracted_values <- MaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```
Retrieving logFC and p values for genes unique to male 
```{r}
#getting male logFC and p values from genes 
MaleUniqueGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

# getting female logFC and p values from genes 
FemaleValuesMaleUniqueGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking if they are identical
identical(MaleUniqueGenes$gene_name, FemaleValuesMaleUniqueGenes$gene_name)
```

Graphing logFC plot 
```{r}
#plotting logFC males and females
plot(MaleUniqueGenes$logFC, FemaleValuesMaleUniqueGenes$logFC)

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
```

```{r}
pdf("LogFCPlotGeneLists/logFcMaleUniqueGenesV3.pdf", width=12, height=12)
#plotting logFC males and females
plot(MaleUniqueGenes$logFC, FemaleValuesMaleUniqueGenes$logFC)

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
dev.off()
```

## **LogFC Plot Genes Unique to Females**

getting genes unique to female
```{r}
# Extract rows where is in female gene list but not male and overall
filtered_rows <- which(!(FemaleGeneList$gene_name %in% MaleGeneList$gene_name | FemaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name))

# Extract the information based on filtered row
extracted_values <- FemaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```
Getting logFC and p values for genes unique to females 
```{r}
#getting female logFC and p values from genes 
FemaleUniqueGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

# getting male logFC and p values from genes 
MaleValuesFemaleUniqueGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking if they are identical
identical(FemaleUniqueGenes$gene_name, MaleValuesFemaleUniqueGenes$gene_name)
```

Graphing logFC plot
```{r}
#plotting logFC males and females
plot(MaleValuesFemaleUniqueGenes$logFC, FemaleUniqueGenes$logFC)

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
```
```{r}
pdf("LogFCPlotGeneLists/logFcFemaleUniqueGenesV3.pdf", width=12, height=12)
#plotting logFC males and females
plot(MaleValuesFemaleUniqueGenes$logFC, FemaleUniqueGenes$logFC)

#adding vertical and horizontal lines 
abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
dev.off()
```

*LogFC Plot Genes Shared between Females and Overall Tumor Tumor-Adjacent*

Getting names of genes shared between females and overall tumor tumor-adjacent
```{r}
# Extract rows where genes are shared between females and overall but not in males
filtered_rows <- which(!(FemaleGeneList$gene_name %in% MaleGeneList$gene_name) & FemaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name)

# Extract the values from the Females column of the filtered rows
extracted_values <- FemaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```

Getting logFC and p values for genes shared between females and overall tumor tumor-adjacent
```{r}
#getting female logFC and p values from genes 
FemaleSharedGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#getting male logFC and p values from genes
MaleValuesFemaleSharedGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking if gene names are identical 
identical(FemaleSharedGenes$gene_name, MaleValuesFemaleSharedGenes$gene_name)
```

graphing logFC plot
```{r}
plot(MaleValuesFemaleSharedGenes$logFC, FemaleSharedGenes$logFC)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
```

```{r}
pdf("LogFCPlotGeneLists/logFcFemaleOverallSharedGenes.pdf", width=12, height=12)
plot(MaleValuesFemaleSharedGenes$logFC, FemaleSharedGenes$logFC)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
dev.off()
```

**LogFC Plot Genes Shared between Overall Tumor Tumor-Adjacent and Males**

Getting genes shared between males and overall tumor tumor-adjacent 
```{r}
# Extract rows where are shared between males and overall but not in females 
filtered_rows <- which(!(MaleGeneList$gene_name %in% FemaleGeneList$gene_name) & MaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name)

# Extract the logFC and p values based on filtered rows
extracted_values <- MaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```

Getting logFC and p values for genes shared between males and overall tumor tumor-adjacent
```{r}
#getting maale logFC and p values from genes 
MaleSharedGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#getting female logFC and p values from genes
FemaleValuesMaleSharedGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking gene names match
identical(MaleSharedGenes$gene_name, FemaleValuesMaleSharedGenes$gene_name)

```

```{r}
plot(MaleSharedGenes$logFC, FemaleValuesMaleSharedGenes$logFC)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
```
```{r}
pdf("LogFCPlotGeneLists/logFcMaleOverallSharedGenesV3.pdf", width=12, height=12)
plot(MaleSharedGenes$logFC, FemaleValuesMaleSharedGenes$logFC)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
dev.off()
```

##*Genes Shared by all three comparisons*
```{r}
filtered_rows <- which(MaleGeneList$gene_name %in% FemaleGeneList$gene_name & MaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name)

# Extract the logFC and p values based on filtered rows
extracted_values <- MaleGeneList$gene_name[filtered_rows]

# Print the extracted values
#print(extracted_values)

filtered_rows2 <- which(FemaleGeneList$gene_name %in% MaleGeneList$gene_name & FemaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name)

# Extract the logFC and p values based on filtered rows
extracted_values2 <- FemaleGeneList$gene_name[filtered_rows2]

# Print the extracted values
#print(extracted_values2)

extracted_values <- extracted_values[order(extracted_values)]

extracted_values2 <- extracted_values2[order(extracted_values2)]

identical(extracted_values, extracted_values2)

filtered_rows3 <- which(OverallGeneList$DEGs.gene_name %in% MaleGeneList$gene_name & OverallGeneList$DEGs.gene_name %in% FemaleGeneList$gene_name)

extracted_values3 <- OverallGeneList$DEGs.gene_name[filtered_rows3]

# Print the extracted values
#print(extracted_values3)

extracted_values3 <- extracted_values3[order(extracted_values3)]

identical(extracted_values, extracted_values3)
```

The genes shared by all three comparisons are equivalent regardless of which gene list you are drawing from. I will be proceeding with only one "extracted value" variable going forward for simplicity. 

```{r}
#getting maale logFC and p values from genes 
MaleOverallSharedGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#getting female logFC and p values from genes
FemaleOverallSharedGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking gene names match
identical(MaleOverallSharedGenes$gene_name, FemaleOverallSharedGenes$gene_name)
```

```{r}
plot(MaleOverallSharedGenes$logFC, FemaleOverallSharedGenes$logFC)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
```

```{r}
pdf("LogFCPlotGeneLists/logFcSharedByAllGenesV3.pdf", width=12, height=12)
plot(MaleOverallSharedGenes$logFC, FemaleOverallSharedGenes$logFC)

abline(v = 0, col = "black", lty = 2)
abline(h = 0, col = "black", lty = 2)
dev.off()
```