---
title: "LogvFoldvChangevPlots"
author: "Annika Jorgensen"
date: "2024-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Libraries, include= FALSE}
#==================================
# Markdown Libraries 
#==================================

#==================================
# DEG Analysis Libraries 
#==================================

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)                                  
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(UpSetR)
library(formatR)
library(tinytex)
```

Importing the filtered and unfiltered gene lists. Filtered lists include genes with logFC greater than or equal to 2 and an adjusted p value of 0.05.
```{r}
#importing differentially expressed gene lists LogFC 2 and p.value 0.05
FemaleGeneList <- read.csv("FemaleTumorTumorAdjacent/Figures/GeneLists/Female_DEGs.csv")
MaleGeneList <- read.csv("MaleTumorTumorAdjacent/Figures/GeneLists/Male_all_DEGs.csv")
OverallGeneList <- read.csv("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/TumorTumorAdjacent/Figures/GeneLists/gene_list_tumor_vs_tumor_adjacent.csv")

#importing unfiltered gene lists
FemaleUnfilteredGeneList <- read.csv("FemaleTumorTumorAdjacent/Figures/GeneLists/unfiltered_DEGs_F.csv")
MaleUnfilteredGeneList <- read.csv("FemaleTumorTumorAdjacent/Figures/GeneLists/unfiltered_DEGs_M.csv")
OverallUnfilteredGeneList <- read.csv("TumorTumorAdjacent/Figures/GeneLists/unfiltered_gene_list_tumor_vs_tumor_adjacent.csv")
```

Creating upset plot from the filtered gene lists
```{r}

#creating upset plot of all differentially expressed genes 
input <- list(OverallGeneList$DEGs.gene_name, MaleGeneList$gene_name, FemaleGeneList$gene_name)

names(input) <- c("all", "male", "female")

upset(fromList(input), order.by= "freq", nsets=3)
```

```{r}
#saving upset plot to PDF
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/DifferentialExpressionAnalysis/upsetplot_sex_stratified.pdf", width = 14, height = 14)
 upset(fromList(input), order.by= "freq", nsets=3, line.size=1.5, text.scale= c(5,5,4.5,2.0,5,5))
 dev.off()
```

Matching the unfiltered gene lists 
```{r}
#matching the male gene list to the female gene list 
FemaleUnfilteredGeneList <- FemaleUnfilteredGeneList[match(MaleUnfilteredGeneList$gene_name, FemaleUnfilteredGeneList$gene_name), ]

#checking if they are identical
identical(FemaleUnfilteredGeneList$gene_name, MaleUnfilteredGeneList$gene_name)
```

Doing regression calculation and creating dataframe for logFC plot
```{r}
#creating regression line
regression <- lm(MaleUnfilteredGeneList$logFC~FemaleUnfilteredGeneList$logFC)

#creating data frame for logFC plot.
df <- data.frame(MaleUnfilteredGeneList$logFC, MaleUnfilteredGeneList$gene_name, MaleUnfilteredGeneList$adj.P.Val, FemaleUnfilteredGeneList$logFC, FemaleUnfilteredGeneList$gene_name, FemaleUnfilteredGeneList$adj.P.Val)

#Placing residuals into data frame for logFC plot
df$residuals <- residuals(regression)

#naming columns 
colnames(df) <- c("M_logFC", "M_name", "M_adj.P.Val", "F_logFC", "F_name", "F_adj.P.Val","residuals")
```

```{r}
#plotting effect size 
ggp<- ggplot(df, aes(x=M_logFC, y= F_logFC))

#adding horizontal and vertical line 
ggp2 <- ggp + geom_abline(intercept=0, slope=0, size=1.5) + # horizontal line 
                 geom_vline(xintercept = 0, size = 1.5) # vertical line 
#adding coloring based on significance
ggp3<- ggp2 + geom_point(alpha = 1.5, aes(color = ifelse(M_adj.P.Val < 0.05 & F_adj.P.Val >= 0.05, "blue", # significant in males
                              ifelse(M_adj.P.Val >= 0.05 & F_adj.P.Val < 0.05, "maroon", # significant in females
                                     ifelse(M_adj.P.Val < 0.05 & F_adj.P.Val < 0.05, "goldenrod", #significant in both
                                     ifelse(M_adj.P.Val >= 0.05 & F_adj.P.Val >=0.05, "black", "green")))))) +
  scale_color_manual(values = c("blue", "maroon","goldenrod", "black"),
                     labels = c("Male Sig., Female Non-Sig.", "Female Sig., Male Non-Sig.","Both Sig", "Both Non-Sig."))
# adding axes labels and legend title 
ggp4 <- ggp3 + labs(color = "Legend") +  # Set legend title using labs()
  theme_bw()

ggp5<- ggp4+theme(axis.title = element_text(family= "Helvetica", size=(35)), axis.text = element_text(family = "Helvetica", size=(35), color="black"), legend.title = element_text(family = "Helvetica", size=(25)), legend.text = element_text(family= "Helvetica", size=(25))) + labs(x= "Male log(FC)", y="Female log(FC)")

ggp5
```

```{r}
pdf("LogFCPlotGeneLists/logFcMaleFemaleColoredBySignificance.pdf", width=12, height=12)
ggp5
dev.off()
```



**LogFC Plot for Genes Unique to Males**
```{r}

# Extract rows where genes are in male list but not female and overall list 
filtered_rows <- which(!(MaleGeneList$gene_name %in% FemaleGeneList$gene_name| MaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name))

# Extract the p value and logFC based on  of the filtered rows
extracted_values <- MaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)


#getting male logFC and p values from genes 
MaleUniqueGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

# getting female logFC and p values from genes 
FemaleValuesMaleUniqueGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#matching the male gene list to the female gene list 
#FemaleUnfilteredGeneList <- FemaleUnfilteredGeneList[match(MaleUnfilteredGeneList$gene_name, FemaleUnfilteredGeneList$gene_name), ]

#checking if they are identical
identical(MaleUniqueGenes$gene_name, FemaleValuesMaleUniqueGenes$gene_name)

#creating data frame with all information 
df <- data.frame(M_logFC = MaleUniqueGenes$logFC, M_name = MaleUniqueGenes$gene_name, M_adj.P.Val = MaleUniqueGenes$adj.P.Val, F_logFC = FemaleValuesMaleUniqueGenes$logFC, F_name = FemaleValuesMaleUniqueGenes$gene_name, F_adj.P.Val = FemaleValuesMaleUniqueGenes$adj.P.Val)
```

```{r}
#plotting effect size 
ggp<- ggplot(df, mapping=aes(x=M_logFC, y= F_logFC))

#adding horizontal and vertical line 
ggp2 <- ggp + geom_abline(intercept=0, slope=0, size=1.5) + # horizontal line 
                 geom_vline(xintercept = 0, size = 1.5) # vertical line 
#adding coloring based on significance
ggp3<- ggp2 + geom_point(alpha = 1.5)

ggp5<- ggp3+theme(axis.title = element_text(family= "Helvetica", size=(35)), axis.text = element_text(family = "Helvetica", size=(35), color="black"), legend.title = element_text(family = "Helvetica", size=(25)), legend.text = element_text(family= "Helvetica", size=(25))) + labs(x= "Male log(FC)", y="Female log(FC)")

ggp5
```

```{r}
pdf("LogFCPlotGeneLists/logFcMaleUniqueGenes.pdf", width=12, height=12)
ggp5
dev.off()
```

**Plotting LogFC for Genes Unique to Females**
```{r}
# Extract rows where is in female gene list but not male and overall
filtered_rows <- which(!(FemaleGeneList$gene_name %in% MaleGeneList$gene_name| FemaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name))

# Extract the information based on filtered row
extracted_values <- FemaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```

```{r}
#getting female logFC and p values from genes 
FemaleUniqueGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

# getting male logFC and p values from genes 
MaleValuesFemaleUniqueGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking if they are identical
identical(FemaleUniqueGenes$gene_name, MaleValuesFemaleUniqueGenes$gene_name)

#building dataframe
df <- data.frame(M_logFC = MaleValuesFemaleUniqueGenes$logFC, M_name = MaleValuesFemaleUniqueGenes$gene_name, M_adj.P.Val = MaleValuesFemaleUniqueGenes$adj.P.Val, F_logFC = FemaleUniqueGenes$logFC, F_name = FemaleUniqueGenes$gene_name, F_adj.P.Val = FemaleUniqueGenes$adj.P.Val)

```

```{r}
#plotting effect size 
ggp<- ggplot(df, mapping=aes(x=M_logFC, y= F_logFC))

#adding horizontal and vertical line 
ggp2 <- ggp + geom_abline(intercept=0, slope=0, size=1.5) + # horizontal line 
                 geom_vline(xintercept = 0, size = 1.5) # vertical line 
#adding coloring based on significance
ggp3<- ggp2 + geom_point(alpha = 1.5)

ggp5<- ggp3+theme(axis.title = element_text(family= "Helvetica", size=(35)), axis.text = element_text(family = "Helvetica", size=(35), color="black"), legend.title = element_text(family = "Helvetica", size=(25)), legend.text = element_text(family= "Helvetica", size=(25))) + labs(x= "Male log(FC)", y="Female log(FC)")

ggp5
```

```{r}
pdf("LogFCPlotGeneLists/logFcFemaleUniqueGenes.pdf", width=12, height=12)
ggp5
dev.off()
```

*LogFC Plot Genes Shared between Females and Overall Tumor Tumor-Adjacent*
```{r}
# Extract rows where genes are shared between females and overall but not in males
filtered_rows <- which(!(FemaleGeneList$gene_name %in% MaleGeneList$gene_name) & FemaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name)

# Extract the values from the Females column of the filtered rows
extracted_values <- FemaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```

```{r}
#getting female logFC and p values from genes 
FemaleSharedGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#getting male logFC and p values from genes
MaleValuesFemaleSharedGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking if gene names are identical 
identical(FemaleSharedGenes$gene_name, MaleValuesFemaleSharedGenes$gene_name)

#building dataframe containing logFC and p values from males and females
df <- data.frame(M_logFC = MaleValuesFemaleSharedGenes$logFC, M_name = MaleValuesFemaleSharedGenes$gene_name, M_adj.P.Val = MaleValuesFemaleSharedGenes$adj.P.Val, F_logFC = FemaleSharedGenes$logFC, F_name = FemaleSharedGenes$gene_name, F_adj.P.Val = FemaleSharedGenes$adj.P.Val)
```

```{r}
#plotting effect size 
ggp<- ggplot(df, mapping=aes(x=M_logFC, y= F_logFC))

#adding horizontal and vertical line 
ggp2 <- ggp + geom_abline(intercept=0, slope=0, size=1.5) + # horizontal line 
                 geom_vline(xintercept = 0, size = 1.5) # vertical line 
#adding coloring based on significance
ggp3<- ggp2 + geom_point(alpha = 1.5)

ggp5<- ggp3+theme(axis.title = element_text(family= "Helvetica", size=(35)), axis.text = element_text(family = "Helvetica", size=(35), color="black"), legend.title = element_text(family = "Helvetica", size=(25)), legend.text = element_text(family= "Helvetica", size=(25))) + labs(x= "Male log(FC)", y="Female log(FC)")

ggp5
```

```{r}
pdf("LogFCPlotGeneLists/logFcFemaleOverallSharedGenes.pdf", width=12, height=12)
ggp5
dev.off()
```

**LogFC Plot Genes Shared between Overall Tumor Tumor-Adjacent and Males**

```{r}
# Extract rows where are shared between males and overall but not in females 
filtered_rows <- which(!(MaleGeneList$gene_name %in% FemaleGeneList$gene_name) & MaleGeneList$gene_name %in% OverallGeneList$DEGs.gene_name)

# Extract the logFC and p values based on filtered rows
extracted_values <- MaleGeneList$gene_name[filtered_rows]

# Print the extracted values
print(extracted_values)
```

```{r}
#getting maale logFC and p values from genes 
MaleSharedGenes <- MaleUnfilteredGeneList[which(MaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#getting female logFC and p values from genes
FemaleValuesMaleSharedGenes <- FemaleUnfilteredGeneList[which(FemaleUnfilteredGeneList$gene_name %in% extracted_values), ]

#checking gene names match
identical(MaleSharedGenes$gene_name, FemaleValuesMaleSharedGenes$gene_name)

#building dataframe of logFc and p values 
df <- data.frame(M_logFC = MaleSharedGenes$logFC, M_name = MaleSharedGenes$gene_name, M_adj.P.Val = MaleSharedGenes$adj.P.Val, F_logFC = FemaleValuesMaleSharedGenes$logFC, F_name = FemaleValuesMaleSharedGenes$gene_name, F_adj.P.Val = FemaleValuesMaleSharedGenes$adj.P.Val)
```

```{r}
#plotting effect size 
ggp<- ggplot(df, mapping=aes(x=M_logFC, y= F_logFC))

#adding horizontal and vertical line 
ggp2 <- ggp + geom_abline(intercept=0, slope=0, size=1.5) + # horizontal line 
                 geom_vline(xintercept = 0, size = 1.5) # vertical line 
#adding coloring based on significance
ggp3<- ggp2 + geom_point(alpha = 1.5)

ggp5<- ggp3+theme(axis.title = element_text(family= "Helvetica", size=(35)), axis.text = element_text(family = "Helvetica", size=(35), color="black"), legend.title = element_text(family = "Helvetica", size=(25)), legend.text = element_text(family= "Helvetica", size=(25))) + labs(x= "Male log(FC)", y="Female log(FC)")

ggp5
```
```{r}
pdf("LogFCPlotGeneLists/logFcMaleOverallSharedGenes.pdf", width=12, height=12)
ggp5
dev.off()
```