---
title: "Survival Analysis Full Data"
output: html_document
date: '12-11-2023'
---

# **Purpose:** 

Perform survival analysis for the full RIKEN cohort males vs. females and subset by etiology and tumor stage


# **Results:**

### Load libraries
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(survminer)
library(survival)
```

### Read in data
```{r}
HCC_metadata <- as.data.frame(read.csv("~/Desktop/ResearchProjects/LiverCancer/Metadata/ICGC_japan_HCC_WGS_samples_info.csv"))
Clinical_metadata <- as.data.frame(read.csv("~/Desktop/ResearchProjects/LiverCancer/Metadata/Full_RIKEN_clinical_data.csv"))

Clinical_metadata$ID <- gsub("_.*", "", Clinical_metadata$ID)
```

### Match data
```{r}
Clinical_metadata <- Clinical_metadata[which(Clinical_metadata$ID %in% HCC_metadata$submitted_donor_id),]
HCC_metadata <- HCC_metadata[which(HCC_metadata$submitted_donor_id %in% Clinical_metadata$ID),]
Clinical_metadata <- Clinical_metadata[match(HCC_metadata$submitted_donor_id, Clinical_metadata$ID),]

identical(HCC_metadata$submitted_donor_id, Clinical_metadata$ID)

Total_metadata <- cbind(HCC_metadata, Clinical_metadata)

#reading in meta dataset
metadataSmall <- read.csv("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/MedicalCovariateAnalysis/data/metadataSubset.csv", row.names=1)
```
```{r}
#pulling sample ids
samples<- as.data.frame(metadataSmall$ID)

# Remove duplicates from the 'sample' column
samples <- unique(samples)
colnames(samples)<- c("RK")

#subsetting survival analysis metadata into samples for analysis
Clinical_metadata <- Clinical_metadata[which(Clinical_metadata$ID %in% samples$RK),]
HCC_metadata<- HCC_metadata[which(HCC_metadata$submitted_donor_id %in% samples$RK),]

#matching data
Clinical_metadata <- Clinical_metadata[match(HCC_metadata$submitted_donor_id, Clinical_metadata$ID),]
identical(HCC_metadata$submitted_donor_id, Clinical_metadata$ID)
Total_metadata <- cbind(HCC_metadata, Clinical_metadata)
```

## Survival analysis all males vs. females: 
```{r}
for (i in 1:length(Total_metadata$donor_vital_status)){
  if (Total_metadata$donor_vital_status[i] == "deceased"){
    Total_metadata$donor_vital_status[i] = 1
  }
  if (Total_metadata$donor_vital_status[i] == "alive"){
    Total_metadata$donor_vital_status[i] =0
  }
}

ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata), pval=TRUE)

pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/SurvivalAnalysis/SurvivalProbabilitybySex.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata), pval=TRUE)
dev.off()
```

## Survival analysis subset males vs. females - HBV: 
```{r}
Total_metadata_HBV <- Total_metadata[which(Total_metadata$Virus.infection=="HBV"),]
  
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HBV), pval=TRUE)

pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/SurvivalAnalysis/SurvivalProbabilitybySexHBVdata.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HBV), pval=TRUE)
dev.off()
```

## Survival analysis subset males vs. females - HCV: 
```{r}
Total_metadata_HCV <- Total_metadata[which(Total_metadata$Virus.infection=="HCV"),]
  
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HCV), pval=TRUE)

pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/SurvivalAnalysis/SurvivalProbabilitybySexHCVdata.pdf")
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_HCV), pval=TRUE)
dev.off()
```
## Survival analysis subset tumor stage :
```{r}
# Find rows where the values in the specified columns are not equal
unequal_rows <- Total_metadata[Total_metadata$donor_tumour_stage_at_diagnosis != Total_metadata$Ta, ]

# Print the rows where the values are not equal
print(unequal_rows)

#11 instances where Ta tumor stage column does not mmatch the tumour stage at diagnosis column
```

```{r}
#using tumor stage at diagnosis
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_tumour_stage_at_diagnosis, data=Total_metadata), pval=TRUE)
```

```{r}
#using Ta
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Ta, data=Total_metadata), pval=TRUE)
```

## Survival analysis subset Stage - Male
```{r}
Total_metadata_male <- Total_metadata[which(Total_metadata$donor_sex=="male"),]

#using tumor stage at diagnosis
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_tumour_stage_at_diagnosis, data=Total_metadata_male), pval=TRUE)
```

```{r }
#using Ta
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Ta, data=Total_metadata_male), pval=TRUE)
```

## Survival analysis subset tumor stage - female
```{r}
Total_metadata_female <- Total_metadata[which(Total_metadata$donor_sex=="female"),]

ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_tumour_stage_at_diagnosis, data=Total_metadata_female), pval=TRUE)
```

```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Ta, data=Total_metadata_female), pval=TRUE)
```

## Survival analysis subset by alcohol 
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Alcohol.intakee, data=Total_metadata), pval=TRUE)
```

## Survival analysis subset alcohol intake - males
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Alcohol.intakee, data=Total_metadata_male), pval=TRUE)
```

## Survival analysis subset alcohol intake - females
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Alcohol.intakee, data=Total_metadata_female), pval=TRUE)
```

## Survival Analysis intakes alcohol versus doesn't intake
```{r}
#changing all values that are not zero to one  reducing the variable to does intake alcohol and doesn't intake
Total_metadata$Alcohol.intakee[Total_metadata$Alcohol.intakee != 0] <- 1

#survival analysis
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Alcohol.intakee, data=Total_metadata), pval=TRUE)
```

## Survival Analysis intakes alcohol versus doesn't intake - males
```{r }
#changing all values that are not zero to one
Total_metadata_male$Alcohol.intakee[Total_metadata_male$Alcohol.intakee != 0] <- 1

#survival analysis
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Alcohol.intakee, data=Total_metadata_male), pval=TRUE)
```

## Survival Analysis intakes alcohol versus doesn't intake - females
```{r }
#changing all values that are not zero to one
Total_metadata_female$Alcohol.intakee[Total_metadata_female$Alcohol.intakee != 0] <- 1

#survival analysis
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Alcohol.intakee, data=Total_metadata_female), pval=TRUE)
```
## Survival analysis subset smoking 
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Smoking, data=Total_metadata), pval=TRUE)
```

## Survival Analysis subset by Smoking - males
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Smoking, data=Total_metadata_male), pval=TRUE)
```

## Survival analysis subset smoking - females
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~Smoking, data=Total_metadata_female), pval=TRUE)
```
## Survival analysis subset viral etiology 

```{r}
# Create a new column 'etiology' based on 'Virus.Infection'
Total_metadata$etiology <- ifelse(Total_metadata$Virus.infection == "HBV", 1, 0)
Total_metadata_female$etiology <- ifelse(Total_metadata_female$Virus.infection == "HBV", 1, 0)
Total_metadata_male$etiology <- ifelse(Total_metadata_male$Virus.infection == "HBV", 1, 0)

ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~etiology, data=Total_metadata), pval=TRUE)
```
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~etiology, data=Total_metadata_male), pval=TRUE)
```
```{r}
ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~etiology, data=Total_metadata_female), pval=TRUE)
```
## Survival analysis subset males vs. females - Neither: 
```{r}
#Total_metadata_NBNC <- Total_metadata[which(Total_metadata$Virus.infection=="NBNC"),]
  
#ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_NBNC), pval=TRUE)

#pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/NBNC_data_survival.pdf")
#ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(donor_vital_status))~donor_sex, data=Total_metadata_NBNC), pval=TRUE)
#dev.off()
```

## Read in French Liver cancer data
```{r}
#data_HCC <- as.data.frame(read.table("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/France Liver Cancer/donor.LICA-FR.tsv",sep="\t", header=TRUE))
#data_HCC <- data_HCC[-which(data_HCC$donor_vital_status==""),]

#for (i in 1:length(data_HCC$donor_vital_status)){
 # if (data_HCC$donor_vital_status[i] == "deceased"){
  #  data_HCC$status[i] = 1
  #}
  #if (data_HCC$donor_vital_status[i] == "alive"){
  #  data_HCC$status[i] =0
  #}
#}

#ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(status))~donor_sex, data=data_HCC), pval=TRUE)

#pdf("C:/Documents/3.0 Hasting Research/Liver Cancer Analysis/Survival/French_data_survival.pdf")
#ggsurvplot(survfit(Surv(donor_survival_time, as.numeric(status))~donor_sex, data=data_HCC), pval=TRUE)
#dev.off()
```

**Interpretation:**

* Across a combination of all of the data, there is significantly worse prognosis for females than males

* In no subset is the difference significant

* There are significant differences in survival based on tumor stage (based on both Ta variable and tumor_stage at diagnosis)

* There are significant differences in survival based on tumor stage in both male and female samples.

* There are no significant differences in survival based on smoking

* There are no significant differences in survival based on Alcohol Intake

**Future steps:**

* Integrate into final figure

