---
title: "X degenerate FPKM expression"
author: "Annika Jorgensen"
date: "2023-10-17"
output: html_document
---

**README**

**Results**
### **Libraries**
The first chunk of code is dedicated to installing the libraries. These libraries are to help execute the differential analysis and helps visualize the data. The code was not included for concision.  
```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(org.Hs.eg.db)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)

setwd('~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples')
```

**Read in data**

This code is where you read in all the data files that are going to be used in the script. The data is also converted into a variety of variables that makes the data easier to handle. The data is also cleaned up to make sure the analysis done later is accurate and precise. 
```{r Reading_in_data}
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",") #changing the name of the file
tumorAdjacentExp <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE) #changing the name of the file 
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp)) #changing the column names 

# Importing gene annotations
#genes <- read.table("gencode.v25.chr_patch_hapl_scaff.annotation.bed", header=FALSE, sep="\t")
genes <- read.table("~/Desktop/ResearchProjects/LiverCancerProject/Metadata/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)

# Removing Samples due to low quality
metadata <- metadata[!(metadata$ID == "RK023") , ]
metadata <- metadata[!(metadata$ID == "RK106") , ]
metadata <- metadata[!(metadata$ID == "RK113") , ]
metadata <- metadata[!(metadata$ID == "RK135") , ]
metadata <- metadata[!(metadata$ID == "RK105") , ]
metadata <- metadata[!(metadata$ID == "RK116") , ]
metadata <- metadata[!(metadata$ID == "RK066") , ]
metadata <- metadata[!(metadata$ID == "RK096") , ]

#Removing both and NBNC samples
metadata <- metadata[!(metadata$Virus_infection == "NBNC"), ]
metadata <- metadata[!(metadata$Virus_infection == "both"), ]

# Subsetting and ordering metadata to match the count matrix
tumorAdjacentExpSubset <- tumorAdjacentExp[,colnames(tumorAdjacentExp) %in% metadata$sampleid]
metadataSubset <- metadata[metadata$sampleid %in% colnames(tumorAdjacentExpSubset),]
metadataSubset <- metadataSubset[match(colnames(tumorAdjacentExpSubset), metadataSubset$sampleid),]
rownames(metadataSubset) <- metadataSubset$sampleid

# Adding tissue type, converting categorical variables to factors
metadataSubset$tumor <- as.numeric(grepl('tumor', metadataSubset$sampleid, ignore.case=T))

#Swapping lesion type for sample RK169
metadataSubset["RK169-tumor-XY","tumor"] <- 0
metadataSubset["RK169-adjacent-XY","tumor"] <- 1

#Changing rownames to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

#Swapping lesion type for sample RK065
metadataSubset["RK065-tumor-XX","tumor"] <- 0
metadataSubset["RK065-adjacent-XX","tumor"] <- 1

#Changing rownames in metadata to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

metadataSubset$gender_tissue <- paste(metadataSubset$Gender, metadataSubset$tumor, sep="_")
metadataSubset$gender_tissue_viral <- paste(metadataSubset$gender_tissue, metadataSubset$Virus_infection, sep="_")
metadataSubset$library_type <- metadataSubset$strandedness
metadataSubset$library_type <- factor(metadataSubset$library_type)
metadataSubset$tumor <- factor(metadataSubset$tumor)
metadataSubset$Ta <- factor(metadataSubset$Ta)
metadataSubset$Portal_vein_invasion <- factor(metadataSubset$Portal_vein_invasion)
metadataSubset$Hepatic_vein_invasion <- factor(metadataSubset$Hepatic_vein_invasion)
metadataSubset$Bile_duct_invasion <- factor(metadataSubset$Bile_duct_invasion)
metadataSubset$Liver_fibrosisc <- factor(metadataSubset$Liver_fibrosisc)
metadataSubset$Prognosis <- factor(metadataSubset$Prognosis)
```

```{r DGE Object}
# Creating the DGEList object
dge <- DGEList(counts=tumorAdjacentExpSubset, genes=genes)
colnames(dge) <- colnames(tumorAdjacentExpSubset)
dge$samples$sex <- metadataSubset$Gender
dge$samples$viral <- factor(metadataSubset$Virus_infection)
dge$samples$ID <- metadataSubset$ID
dge$samples$tumor <- metadataSubset$tumor
dge$samples$gender_tissue <- metadataSubset$gender_tissue
dge$samples$gender_tissue_viral <- metadataSubset$gender_tissue_viral
dge$samples$library_type <- metadataSubset$library_type
dge$samples$edmonson_grade <- metadataSubset$Edmondson_grade
dge$samples$Ta <- metadataSubset$Ta
dge$samples$survival <- metadataSubset$Overall_survival_month

# Inspecting the N of samples in each group
table(dge$samples$gender_tissue_viral)
```

```{r fpkm}
# ======================================
# Filtering expression data
# ======================================

# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the groups under investigation 
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]

# N of genes retained after filtering
dim(dge$genes)
```

```{r take three}

```
**Scraps**
```{r pulling out genes}

# Pulling X degenerate expression data from dge variable into individual variables

dge <- as.data.frame(dge)
gene0 <- dge[which(dge$gene_name=="KDM5D"),]
gene1 <- dge[which(dge$gene_name=="DDX3Y"),]
gene2 <- dge[which(dge$gene_name=="EIF1AY"),]
gene3 <- dge[which(dge$gene_name=="TXLNGY"),]
gene4 <- dge[which(dge$gene_name=="TSPY1"),]
gene5 <- dge[which(dge$gene_name=="AR"),]
gene6 <- dge[which(dge$gene_name=="XIST"),]

#creating data frame of all genes we will be working with
xDegenerateExpression<- rbind(gene0,gene1,gene2,gene3,gene4)
rownames(xDegenerateExpression) <- xDegenerateExpression$gene_name
xDegenerateExpression <- subset(xDegenerateExpression, select = -c(chr, start, end, TXNAME, GENEID, length,gene_name))
xDegenerateExpression<- t(xDegenerateExpression)
xDegenerateExpression <- as.data.frame(xDegenerateExpression)
#making the gene names the rownames 
#rownames(xDegenerateExpression) <- xDegenerateExpression$gene_name

#subsetting out all columns that are not expression data
 # xDegenerateExpression <- as.data.frame(t(xDegenerateExpression))

#subsetting by tumor 
Tumor_expressionData <- xDegenerateExpression[grepl("tumor",rownames(xDegenerateExpression),fixed=TRUE),]
Tumor_expressionData <- as.data.frame(Tumor_expressionData)
Tumor_expressionData<- t(Tumor_expressionData)
Tumor_expressionData <- as.data.frame(Tumor_expressionData)

Adjacent_expressionData <- xDegenerateExpression[grepl("adjacent",rownames(xDegenerateExpression),fixed=TRUE),]
Adjacent_expressionData <- as.data.frame(Adjacent_expressionData)
Adjacent_expressionData<- t(Adjacent_expressionData)
Adjacent_expressionData <- as.data.frame(Adjacent_expressionData)

#Creating Names column
#for (i in 1:dim(Tumor_expressionData)[1]) {
#    Tumor_expressionData$Names[i] = substr(row.names(Tumor_expressionData)[i],3,5)
#}
#Tumor_expressionData$group<- 1:149
```

```{r test}
#test <- Tumor_expressionData[1,]
#test <- t(test)
#test <- as.data.frame(test)
#test$gene <- rownames(test)
#colnames(test) <- c("RK002","gene")
#ggplot(test, aes(x=gene, y=RK002)) + geom_line() + geom_point()
#ggplot(data=Tumor_expressionData,aes(x=gene, y=Names, group=group)) +geom_line() + geom_point()
```

```{r take two}




```

```{r scatterplot}
geneXaxis<- 1:5

pdf("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/newFigures/SamplesXdegenerateGenesPlot.pdf", width=12, height=12)
plot(geneXaxis,Tumor_expressionData$`RK002-tumor-XY`,main="FPKM of all Tumor Samples for KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1",type="b",ylab="FPKM",xlab="KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1")

for (i in colnames(Tumor_expressionData)) {
  lines(geneXaxis,Tumor_expressionData[,i],type="b")
  text(geneXaxis, Tumor_expressionData[,i],
     labels = colnames(Tumor_expressionData),
     cex = 0.6, pos = 4, col = "red",)
}
dev.off()

pdf("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/newFigures/SamplesXdegenerateGenesPlot_adjacent.pdf", width=12, height=12)
plot(geneXaxis,Adjacent_expressionData$`RK002-adjacent-XY`,main="FPKM of all Tumor Samples for KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1",type="b",ylab="FPKM",xlab="KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1")

for (i in colnames(Adjacent_expressionData)) {
  lines(geneXaxis,Adjacent_expressionData[,i],type="b")
  text(geneXaxis, Adjacent_expressionData[,i],
     labels = colnames(Adjacent_expressionData),
     cex = 0.6, pos = 4, col = "red",)
}
dev.off()
#for( i in colnames(Tumor_expressionData)) {
  
  #plot(geneXaxis, Tumor_expressionData[,i],type="b")
  
  
  
#}
pdf("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/newFigures/SamplesXdegenerateGenesPlotAxisLim.pdf", width=12, height=12)
plot(geneXaxis,Tumor_expressionData$`RK002-tumor-XY`,main="FPKM of all Tumor Samples for KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1",type="b",ylab="FPKM",xlab="KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1", ylim=c(0,100))
for (i in colnames(Tumor_expressionData)) {
  lines(geneXaxis,Tumor_expressionData[,i],type="b")
  text(geneXaxis, Tumor_expressionData[,i],
    labels = colnames(Tumor_expressionData),
     cex = 1, pos = 4,col = "red",adj=0.5)
}
dev.off()

pdf("~/Desktop/ResearchProjects/LiverCancerProject/DEA_removed_samples/newFigures/SamplesXdegenerateGenesPlotAxisLim_adjacent.pdf", width=12, height=12)
plot(geneXaxis,Adjacent_expressionData$`RK002-adjacent-XY`,main="FPKM of all Tumor Samples for KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1",type="b",ylab="FPKM",xlab="KDM5D, DDX3Y, EIF1AY, TXLNGY, TSPY1", ylim=c(0,100))

for (i in colnames(Adjacent_expressionData)) {
  lines(geneXaxis,Adjacent_expressionData[,i],type="b")
  text(geneXaxis, Adjacent_expressionData[,i],
     labels = colnames(Adjacent_expressionData),
     cex = 0.6, pos = 4, col = "red",)
}
dev.off()
#creating scatterplot
#ggplot(gene0,aes(x=, y=Expression)) + geom_point()
```
