---
title: "2023_06_07_v2_MDS_plots"
author: "Annika Jorgensen"
date: "2023-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **R Markdown**

Title: "2023 06/07 Version 2 New MDS Plots"  
Author: Annika Jorgensen  
Date: 06/07/2023 

**Purpose**
This script is meant to remove the samples that were found to be outliers and then remake the MDS plot 

**Methods**


**Results**
### **Libraries**
The first chunk of code is dedicated to installing the libraries. These libraries are to help execute the differential analysis and helps visualize the data. The code was not included for concision.  
```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)

setwd('~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples')
```

**Defining Colors** 

This chunk defines color palette variables that are going to be used in plots later on the script. These variables are defined by converting BrewerCode palettes into palettes that can be used in R. 
```{r Color_Palettes}
viralPalette <- brewer.pal(8, "Set1")
hbvColor <- viralPalette[1]
hcvColor <- viralPalette[2]
bothColor <- viralPalette[3]
neitherColor <- viralPalette[4]

sexTissuePalette <- brewer.pal(12, "Paired")
maleTumorColor <- sexTissuePalette[4]
maleAdjacentColor <- sexTissuePalette[3]
femaleTumorColor <- sexTissuePalette[6]
femaleAdjacentColor <- sexTissuePalette[5]
```

**Read in data**

This code is where you read in all the data files that are going to be used in the script. The data is also converted into a variety of variables that makes the data easier to handle. The data is also cleaned up to make sure the analysis done later is accurate and precise. 
```{r Reading_in_data}
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",") #changing the name of the file
tumorAdjacentExp <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE) #changing the name of the file 
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp)) #changing the column names 

# Importing gene annotations
#genes <- read.table("gencode.v25.chr_patch_hapl_scaff.annotation.bed", header=FALSE, sep="\t")
genes <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)

# Removing Samples due to low quality
metadata <- metadata[!(metadata$ID == "RK023") , ]
metadata <- metadata[!(metadata$ID == "RK106") , ]
metadata <- metadata[!(metadata$ID == "RK113") , ]
metadata <- metadata[!(metadata$ID == "RK135") , ]
metadata <- metadata[!(metadata$ID == "RK105") , ]
metadata <- metadata[!(metadata$ID == "RK116") , ]
metadata <- metadata[!(metadata$ID == "RK066") , ]
metadata <- metadata[!(metadata$ID == "RK096") , ]

#Removing both and NBNC samples
metadata <- metadata[!(metadata$Virus_infection == "NBNC"), ]
metadata <- metadata[!(metadata$Virus_infection == "both"), ]

# Subsetting and ordering metadata to match the count matrix
tumorAdjacentExpSubset <- tumorAdjacentExp[,colnames(tumorAdjacentExp) %in% metadata$sampleid]
metadataSubset <- metadata[metadata$sampleid %in% colnames(tumorAdjacentExpSubset),]
metadataSubset <- metadataSubset[match(colnames(tumorAdjacentExpSubset), metadataSubset$sampleid),]
rownames(metadataSubset) <- metadataSubset$sampleid

# Adding tissue type, converting categorical variables to factors
metadataSubset$tumor <- as.numeric(grepl('tumor', metadataSubset$sampleid, ignore.case=T))

#Swapping lesion type for sample RK169
metadataSubset["RK169-tumor-XY","tumor"] <- 0
metadataSubset["RK169-adjacent-XY","tumor"] <- 1

#Changing rownames to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

#Swapping lesion type for sample RK065
metadataSubset["RK065-tumor-XX","tumor"] <- 0
metadataSubset["RK065-adjacent-XX","tumor"] <- 1

#Changing rownames in metadata to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

metadataSubset$gender_tissue <- paste(metadataSubset$Gender, metadataSubset$tumor, sep="_")
metadataSubset$gender_tissue_viral <- paste(metadataSubset$gender_tissue, metadataSubset$Virus_infection, sep="_")
metadataSubset$library_type <- metadataSubset$strandedness
metadataSubset$library_type <- factor(metadataSubset$library_type)
metadataSubset$tumor <- factor(metadataSubset$tumor)
metadataSubset$Ta <- factor(metadataSubset$Ta)
metadataSubset$Portal_vein_invasion <- factor(metadataSubset$Portal_vein_invasion)
metadataSubset$Hepatic_vein_invasion <- factor(metadataSubset$Hepatic_vein_invasion)
metadataSubset$Bile_duct_invasion <- factor(metadataSubset$Bile_duct_invasion)
metadataSubset$Liver_fibrosisc <- factor(metadataSubset$Liver_fibrosisc)
metadataSubset$Prognosis <- factor(metadataSubset$Prognosis)
```

```{r DGE Object}
# Creating the DGEList object
dge <- DGEList(counts=tumorAdjacentExpSubset, genes=genes)
colnames(dge) <- colnames(tumorAdjacentExpSubset)
dge$samples$sex <- metadataSubset$Gender
dge$samples$viral <- factor(metadataSubset$Virus_infection)
dge$samples$ID <- metadataSubset$ID
dge$samples$tumor <- metadataSubset$tumor
dge$samples$gender_tissue <- metadataSubset$gender_tissue
dge$samples$gender_tissue_viral <- metadataSubset$gender_tissue_viral
dge$samples$library_type <- metadataSubset$library_type
dge$samples$edmonson_grade <- metadataSubset$Edmondson_grade
dge$samples$Ta <- metadataSubset$Ta
dge$samples$survival <- metadataSubset$Overall_survival_month
dge$samples$smoking <- factor(metadataSubset$Smoking)
dge$samples$alcohol <- factor(metadataSubset$Alcohol_intake)
dge$samples$fibrosis <- factor(metadataSubset$Liver_fibrosisc)

# Inspecting the N of samples in each group
table(dge$samples$gender_tissue_viral)
```

```{r fpkm}
# ======================================
# Filtering expression data
# ======================================

# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

```{r Voom}
# ===================================== ======================
# ===========================================================
# Analysis of all tumor vs. tumor-adjacent regardless of sex
# ===========================================================
# ===========================================================

# Creating a new design model matrix with the variable of interest and the
# library type
design <- model.matrix(~0+dge$samples$tumor+dge$samples$library_type+dge$samples$Ta+dge$samples$smoking+dge$samples$alcohol+dge$samples$fibrosis)
colnames(design) <- gsub("dge\\$samples\\$tumor", "tumor", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$Ta2", "Ta2", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$Ta3", "Ta3", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$Ta4", "Ta4", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$smoking", "smoking", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$alcohol", "alcohol", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$fibrosis", "fibrosis", colnames(design))
#colnames(design) <- gsub("dge\\$samples\\$Ta4", "Ta4", colnames(design))

head(design)

# Running voom again with the new design matrix.
v <- voomWithQualityWeights(dge, design, plot=TRUE)

```

**Top 25 gene MDS Plot**
This creates outputs an MDS Plot that we've previously see showing the dissimilarity between Male and Female Tumor-Tumor adjacent 

```{r Library Type Correction}
# Removing batch effects
vCorrectLibtype <- removeBatchEffect(v, batch=v$targets$library_type)
write.csv(vCorrectLibtype,"~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/data/MDSplotData.csv")
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/removed_samples_MDS_plots_top_25_genes.pdf")
mds <- plotMDS(vCorrectLibtype, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(v$targets$viral %in% c("HBV"), 17,
                          ifelse(v$targets$viral %in% c("HCV"), 15,
                                 ifelse(v$targets$viral %in% c("both"), 16,  3))),
               col=ifelse(v$targets$gender_tissue=="M_1",
                          "#FFC20A",
                          ifelse(v$targets$gender_tissue=="M_0",
                                 "#E66100",
                                 ifelse(v$targets$gender_tissue=="F_1",
                                        "#40B0A6", "#006CD1"))),
               gene.selection = "common")
legend("center", pch=c(15), 
       col=c( "#FFC20A", "#40B0A6","#E66100", "#006CD1"),
      legend=c("Male tumor","Female tumor", "Male adjacent","Female adjacent"))
legend("topright", pch=c(17, 15, 16, 3),
       col=c("black"),
       legend=c("HBV", "HCV"))
dev.off()

mds <- plotMDS(vCorrectLibtype, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(v$targets$viral %in% c("HBV"), 17,
                          ifelse(v$targets$viral %in% c("HCV"), 15,
                                ifelse(v$targets$viral %in% c("both"), 16,  3))),
               col=ifelse(v$targets$gender_tissue=="M_1",
                          "#FFC20A",
                         ifelse(v$targets$gender_tissue=="M_0",
                                "#E66100",
                                 ifelse(v$targets$gender_tissue=="F_1",
                                        "#40B0A6", "#006CD1"))),
               gene.selection = "common")
legend("center", pch=c(15), 
       col=c( "#FFC20A", "#40B0A6","#E66100", "#006CD1"),
      legend=c("Male tumor","Female tumor", "Male adjacent","Female adjacent"))
legend("topright", pch=c(17, 15, 16, 3),
 #     col=c("black"),
       legend=c("HBV", "HCV"))
```

```{r Tumor Samples}
#creating column with stage and sex information 
viral<- as.data.frame(v[["targets"]][["viral"]])
identifiers<- cbind(stage=v[["targets"]][["Ta"]],sex=v[["targets"]][["sex"]], viral)

#getting expression data from voom output
samples<- as.data.frame(v)

#removing columns that do not contain expression data
samples<- subset(samples, select = -c(chr, start, end, TXNAME, GENEID, length,gene_name))

#rownames of stage column contain samples names
rownames(identifiers)<- colnames(samples)


#subsetting by tumor samples
identifiers<-identifiers[grepl("tumor",rownames(identifiers),fixed=TRUE),]
identifiers<- as.data.frame(identifiers)

# checking number of female and male tumor samples match 
Tumor<- vCorrectLibtype[, grepl("tumor",colnames(vCorrectLibtype),fixed=TRUE)]
Tumor<- as.data.frame(Tumor)

are_rownames_matching <- identical(rownames(identifiers), colnames(Tumor))
print(are_rownames_matching)

```

```{r tumor sample variation}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/removed_samples_MDS_plots_tumor_top_25_genes.pdf")
mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$stage=="1","#FFC20A",
                           ifelse(identifiers$stage=="2", "#E66100",
                                 ifelse(identifiers$stage=="3", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("Stage 1","Stage 2", "Stage 3","Stage 4"))
legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  

```

```{r tumor sample variation}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/removed_samples_MDS_plots_tumor_top_50_genes.pdf")
mds <- plotMDS(Tumor, top = 50, ndim = 10, dim.plot = c(3,4), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$stage=="1","#FFC20A",
                           ifelse(identifiers$stage=="2", "#E66100",
                                 ifelse(identifiers$stage=="3", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("Stage 1","Stage 2", "Stage 3","Stage 4")) 

legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  
```

```{r tumor sample variation}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/removed_samples_MDS_plots_tumor_top_100_genes.pdf")
mds <- plotMDS(Tumor, top = 100, ndim = 10, dim.plot = c(3,4), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$stage=="1","#FFC20A",
                           ifelse(identifiers$stage=="2", "#E66100",
                                 ifelse(identifiers$stage=="3", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("Stage 1","Stage 2", "Stage 3","Stage 4")) 

legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))           
```

```{r tumor sample variation}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/removed_samples_MDS_plots_tumor_top_1000_genes.pdf")
mds <- plotMDS(Tumor, top = 1000, ndim = 10, dim.plot = c(3,4), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$stage=="1","#FFC20A",
                           ifelse(identifiers$stage=="2", "#E66100",
                                 ifelse(identifiers$stage=="3", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("Stage 1","Stage 2", "Stage 3","Stage 4")) 

legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))         
```

```{r adding smoking}
tumorMetadataSubset <- metadataSubset[grepl("tumor",rownames(metadataSubset),fixed=TRUE),]
are_rownames_matching <- identical(rownames(identifiers), rownames(tumorMetadataSubset))
print(are_rownames_matching)

tumorMetadataSubset <- tumorMetadataSubset[match(rownames(tumorMetadataSubset), rownames(identifiers)),]
are_rownames_matching <- identical(rownames(tumorMetadataSubset), rownames(identifiers))
print(are_rownames_matching)


identifiers<- cbind(identifiers,tumorMetadataSubset$Edmondson_grade)
colnames(identifiers)<- c("stage","sex","viral","grade")
```

```{r virus}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/virus.pdf")
plotMDS(Tumor, top = 1000, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col=ifelse(identifiers$viral %in% c("HBV"), "#1A85FF", 
                          ifelse(identifiers$viral %in% c("HCV"), "#D41159", "black")),
               gene.selection = "common")
legend("top", pch=c(15), 
       col=c("#D41159", "#1A85FF") ,
       legend=c("HCV", "HBV"))
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
```

```{r edmondson grade}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/testing.pdf")
plotMDS(Tumor, top = 25, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
              col= ifelse(identifiers$grade=="0","red",
                           ifelse(identifiers$grade=="1", "orange",
                                 ifelse(identifiers$grade=="1~2", "yellow",
                                      ifelse(identifiers$grade=="2",  "green",
                                             ifelse(identifiers$grade=="2~3", "darkgreen",
                                                    ifelse(identifiers$grade=="3", "blue", "violet")))))),
               gene.selection = "common")
legend("top", pch=c(15), 
       col=c( "red", "orange","yellow", "green", "darkgreen", "blue", "violet"),
      legend=c("Grade 0","Grade 1", "Grade 1.5","Grade 2", "Grade 2.5", "Grade 3", "Grade 4")) 
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
```

```{r smoking}
identifiers<- cbind(identifiers,tumorMetadataSubset$Smoking)
colnames(identifiers)<- c("stage","sex","viral","grade","Smoking")
```
```{r plot smoking}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/smoking.pdf")
plotMDS(Tumor, top = 25, dim.plot = c(3,4), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col=ifelse(identifiers$Smoking %in% c("1"), "#1A85FF", "#D41159"),
               gene.selection = "common")
legend("bottomright", pch=c(15), 
       col=c("#D41159", "#1A85FF") ,
       legend=c("Not smoking", "Smoking"))
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
dev.off()

plotMDS(Tumor, top = 25, dim.plot = c(3,4), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col=ifelse(identifiers$Smoking %in% c("1"), "#1A85FF", "#D41159"),
               gene.selection = "common")
legend("topleft", pch=c(15), 
       col=c("#D41159", "#1A85FF") ,
       legend=c("Not smoking", "Smoking"))
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
```

```{r alcohol}
identifiers<- cbind(identifiers,tumorMetadataSubset$Alcohol_intake)
colnames(identifiers)<- c("stage","sex","viral","grade","Smoking", "Alcohol")
```

```{r alcohol mds plot}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/alcohol.pdf")
mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$Alcohol=="0","#FFC20A",
                           ifelse(identifiers$Alcohol=="1", "#E66100",
                                 ifelse(identifiers$Alcohol=="2", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottomright", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("No alcohol","Severity 1", "Severity 2","Severity 3")) 

legend("topleft", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male")) 
dev.off()

mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$Alcohol=="0","#FFC20A",
                           ifelse(identifiers$Alcohol=="1", "#E66100",
                                 ifelse(identifiers$Alcohol=="2", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("No alcohol","Severity 1", "Severity 2","Severity 3")) 

legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  
```

```{r fibrosis}
identifiers<- cbind(identifiers,tumorMetadataSubset$Liver_fibrosisc)
colnames(identifiers)<- c("stage","sex","viral","grade","Smoking", "Alcohol", "Fibrosis")
```

```{r plot fibrosis}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/fibrosis.pdf")
mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$Fibrosis=="0","#FFC20A",
                           ifelse(identifiers$Fibrosis=="1", "#E66100",
                                 ifelse(identifiers$Fibrosis=="2", "#40B0A6",
                                        ifelse(identifiers$Fibrosis== "3", "#006CD1", "darkblue")))), gene.selection = "common")

legend("center", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1", "darkblue"),
      legend=c("No Fibrosis","Severity 1", "Severity 2","Severity 3", "Severity 4")) 

legend("topleft", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  
dev.off()

mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$Fibrosis=="0","#FFC20A",
                           ifelse(identifiers$Fibrosis=="1", "#E66100",
                                 ifelse(identifiers$Fibrosis=="2", "#40B0A6",
                                        ifelse(identifiers$Fibrosis== "3", "#006CD1", "darkblue")))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1", "darkblue"),
      legend=c("No Fibrosis","Severity 1", "Severity 2","Severity 3", "Severity 4")) 

legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  
```