---
title: "2023_06_07_v2_MDS_plots"
author: "Annika Jorgensen"
date: "2023-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **R Markdown**

Title: "2023 06/07 Version 2 New MDS Plots"  
Author: Annika Jorgensen  
Date: 06/07/2023 

**Purpose**
The purpose of this file is twofold. This script is meant to remove the samples that were found to be low quality and then remake the MDS plot to illustrate the low quality samples were removed. This script also conducts MDS analysis on a number of variables in the metadata thought to be significantly different between males and females to see if the variables are potentially driving variation in the tumor samples.

**Results**
### **Libraries**
The first chunk of code is dedicated to installing the libraries. These libraries are to help execute the differential analysis and helps visualize the data. The code was not included for concision.  
```{r Libraries, include=FALSE}

library(RColorBrewer)
library(matrixStats)
library(ggplot2)
library(edgeR)
library(DESeq2)
library(limma)
library(doParallel)
library(variancePartition)
library(clusterProfiler)
library(GOSemSim)
library(biomaRt)
library(UpSetR)
library(VennDiagram)
library(ggrepel)
library(dplyr)
library(stringr)
library(forcats)
library(stats)
library(gridExtra)

setwd('~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples')
```

**Defining Colors** 

This chunk defines color palette variables that are going to be used in plots later on the script. These variables are defined by converting BrewerCode palettes into palettes that can be used in R. 
```{r Color_Palettes}
viralPalette <- brewer.pal(8, "Set1")
hbvColor <- viralPalette[1]
hcvColor <- viralPalette[2]
bothColor <- viralPalette[3]
neitherColor <- viralPalette[4]

sexTissuePalette <- brewer.pal(12, "Paired")
maleTumorColor <- sexTissuePalette[4]
maleAdjacentColor <- sexTissuePalette[3]
femaleTumorColor <- sexTissuePalette[6]
femaleAdjacentColor <- sexTissuePalette[5]
```

**Read in data**

This code is where you read in all the data files that are going to be used in the script. The data is also converted into a variety of variables that makes the data easier to handle. The data is also cleaned up to make sure the analysis done later is accurate and precise. 
```{r Reading_in_data}
metadata <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",") #changing the name of the file
tumorAdjacentExp <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE) #changing the name of the file 
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp)) #changing the column names 

# Importing gene annotations
#genes <- read.table("gencode.v25.chr_patch_hapl_scaff.annotation.bed", header=FALSE, sep="\t")
genes <- read.table("~/Desktop/ResearchProjects/LiverCancer/Metadata/gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)

# Removing Samples due to low quality
metadata <- metadata[!(metadata$ID == "RK023") , ]
metadata <- metadata[!(metadata$ID == "RK106") , ]
metadata <- metadata[!(metadata$ID == "RK113") , ]
metadata <- metadata[!(metadata$ID == "RK135") , ]
metadata <- metadata[!(metadata$ID == "RK105") , ]
metadata <- metadata[!(metadata$ID == "RK116") , ]
metadata <- metadata[!(metadata$ID == "RK066") , ]
metadata <- metadata[!(metadata$ID == "RK096") , ]

#Removing both and NBNC samples
metadata <- metadata[!(metadata$Virus_infection == "NBNC"), ]
metadata <- metadata[!(metadata$Virus_infection == "both"), ]

# Subsetting and ordering metadata to match the count matrix
tumorAdjacentExpSubset <- tumorAdjacentExp[,colnames(tumorAdjacentExp) %in% metadata$sampleid]
metadataSubset <- metadata[metadata$sampleid %in% colnames(tumorAdjacentExpSubset),]
metadataSubset <- metadataSubset[match(colnames(tumorAdjacentExpSubset), metadataSubset$sampleid),]
rownames(metadataSubset) <- metadataSubset$sampleid

# Adding tissue type, converting categorical variables to factors
metadataSubset$tumor <- as.numeric(grepl('tumor', metadataSubset$sampleid, ignore.case=T))

#Swapping lesion type for sample RK169
metadataSubset["RK169-tumor-XY","tumor"] <- 0
metadataSubset["RK169-adjacent-XY","tumor"] <- 1

#Changing rownames to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-tumor-XY"] <- "RK169_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169-adjacent-XY"] <- "RK169_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK169_adjacent-XY"] <- "RK169-adjacent-XY"

#Swapping lesion type for sample RK065
metadataSubset["RK065-tumor-XX","tumor"] <- 0
metadataSubset["RK065-adjacent-XX","tumor"] <- 1

#Changing rownames in metadata to match swapped lesion type
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(metadataSubset)[rownames(metadataSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-tumor-XY"] <- "RK065_adjacent-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065-adjacent-XY"] <- "RK065_tumor-XY"
rownames(tumorAdjacentExpSubset)[rownames(tumorAdjacentExpSubset)=="RK065_adjacent-XY"] <- "RK065-adjacent-XY"

metadataSubset$gender_tissue <- paste(metadataSubset$Gender, metadataSubset$tumor, sep="_")
metadataSubset$gender_tissue_viral <- paste(metadataSubset$gender_tissue, metadataSubset$Virus_infection, sep="_")
metadataSubset$library_type <- metadataSubset$strandedness
metadataSubset$library_type <- factor(metadataSubset$library_type)
metadataSubset$tumor <- factor(metadataSubset$tumor)
metadataSubset$Ta <- factor(metadataSubset$Ta)
metadataSubset$Portal_vein_invasion <- factor(metadataSubset$Portal_vein_invasion)
metadataSubset$Hepatic_vein_invasion <- factor(metadataSubset$Hepatic_vein_invasion)
metadataSubset$Bile_duct_invasion <- factor(metadataSubset$Bile_duct_invasion)
metadataSubset$Liver_fibrosisc <- factor(metadataSubset$Liver_fibrosisc)
metadataSubset$Prognosis <- factor(metadataSubset$Prognosis)
```

```{r DGE Object}
# Creating the DGEList object
dge <- DGEList(counts=tumorAdjacentExpSubset, genes=genes)
colnames(dge) <- colnames(tumorAdjacentExpSubset)
dge$samples$sex <- metadataSubset$Gender
dge$samples$viral <- factor(metadataSubset$Virus_infection)
dge$samples$ID <- metadataSubset$ID
dge$samples$tumor <- metadataSubset$tumor
dge$samples$gender_tissue <- metadataSubset$gender_tissue
dge$samples$gender_tissue_viral <- metadataSubset$gender_tissue_viral
dge$samples$library_type <- metadataSubset$library_type
dge$samples$edmonson_grade <- metadataSubset$Edmondson_grade
dge$samples$Ta <- metadataSubset$Ta
dge$samples$survival <- metadataSubset$Overall_survival_month
dge$samples$smoking <- factor(metadataSubset$Smoking)
dge$samples$alcohol <- factor(metadataSubset$Alcohol_intake)
dge$samples$fibrosis <- factor(metadataSubset$Liver_fibrosisc)

# Inspecting the N of samples in each group
table(dge$samples$gender_tissue_viral)
```

```{r fpkm}
# ======================================
# Filtering expression data
# ======================================

# Keeping genes that have a mean FPKM of at least 0.5 in at least one of the
# groups under investigation and at least 6 reads in at least 10 samples
fpkm <- rpkm(dge, gene.length=dge$genes$length)

M_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HBV")],1,mean,na.rm=TRUE)
M_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HBV")],1,mean,na.rm=TRUE)
M_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_1_HCV")],1,mean,na.rm=TRUE)
M_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="M_0_HCV")],1,mean,na.rm=TRUE)

F_1_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HBV")],1,mean,na.rm=TRUE)
F_0_HBV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HBV")],1,mean,na.rm=TRUE)
F_1_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_1_HCV")],1,mean,na.rm=TRUE)
F_0_HCV_mean_fpkm <- apply(as.data.frame(fpkm)[(dge$samples$gender_tissue_viral=="F_0_HCV")],1,mean,na.rm=TRUE)

keep <- (M_1_HBV_mean_fpkm > 0.5 | M_0_HBV_mean_fpkm > 0.5 | 
           M_1_HCV_mean_fpkm > 0.5 | M_0_HCV_mean_fpkm > 0.5 |
           F_1_HBV_mean_fpkm > 0.5 | F_0_HBV_mean_fpkm > 0.5 |
           F_1_HCV_mean_fpkm > 0.5 | F_0_HCV_mean_fpkm > 0.5 )

dge <- dge[keep,,keep.lib.sizes=FALSE]
dge <- calcNormFactors(dge, method="TMM")
keep <- rowSums(dge$counts > 6) >= 10
dge <- dge[keep,,keep.lib.size=FALSE]
dge <- calcNormFactors(dge, method="TMM")

# N of genes retained after filtering
dim(dge$genes)
```

```{r Voom}
# ===================================== ======================
# ===========================================================
# Analysis of all tumor vs. tumor-adjacent regardless of sex
# ===========================================================
# ===========================================================

# Creating a new design model matrix with the variable of interest and the
# library type
design <- model.matrix(~0+dge$samples$tumor+dge$samples$library_type+dge$samples$Ta+dge$samples$smoking+dge$samples$alcohol+dge$samples$fibrosis)
colnames(design) <- gsub("dge\\$samples\\$tumor", "tumor", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$library_typeunstranded", "library_type", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$Ta2", "Ta2", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$Ta3", "Ta3", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$Ta4", "Ta4", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$smoking", "smoking", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$alcohol", "alcohol", colnames(design))
colnames(design) <- gsub("dge\\$samples\\$fibrosis", "fibrosis", colnames(design))
#colnames(design) <- gsub("dge\\$samples\\$Ta4", "Ta4", colnames(design))

head(design)

# Running voom again with the new design matrix.
v <- voomWithQualityWeights(dge, design, plot=TRUE)

```

**Top 25 gene MDS Plot**

Recreating the MDS plot we have seen previously with the removed samples to see if all low quality samples have been removed.
```{r Library Type Correction}
# Removing batch effects
vCorrectLibtype <- removeBatchEffect(v, batch=v$targets$library_type)

#creating MDS plot top 25 genes PC 1 and 2 coloring by lesion type and sex
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/removed_samples_MDS_plots_top_25_genes.pdf")
mds <- plotMDS(vCorrectLibtype, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(v$targets$viral %in% c("HBV"), 17,
                          ifelse(v$targets$viral %in% c("HCV"), 15,
                                 ifelse(v$targets$viral %in% c("both"), 16,  3))),
               col=ifelse(v$targets$gender_tissue=="M_1",
                          "#FFC20A",
                          ifelse(v$targets$gender_tissue=="M_0",
                                 "#E66100",
                                 ifelse(v$targets$gender_tissue=="F_1",
                                        "#40B0A6", "#006CD1"))),
               gene.selection = "common")
legend("center", pch=c(15), 
       col=c( "#FFC20A", "#40B0A6","#E66100", "#006CD1"),
      legend=c("Male tumor","Female tumor", "Male adjacent","Female adjacent"))
legend("topright", pch=c(17, 15, 16, 3),
       col=c("black"),
       legend=c("HBV", "HCV"))
dev.off()

mds <- plotMDS(vCorrectLibtype, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(v$targets$viral %in% c("HBV"), 17,
                          ifelse(v$targets$viral %in% c("HCV"), 15,
                                ifelse(v$targets$viral %in% c("both"), 16,  3))),
               col=ifelse(v$targets$gender_tissue=="M_1",
                          "#FFC20A",
                         ifelse(v$targets$gender_tissue=="M_0",
                                "#E66100",
                                 ifelse(v$targets$gender_tissue=="F_1",
                                        "#40B0A6", "#006CD1"))),
               gene.selection = "common")
legend("center", pch=c(15), 
       col=c( "#FFC20A", "#40B0A6","#E66100", "#006CD1"),
      legend=c("Male tumor","Female tumor", "Male adjacent","Female adjacent"))
legend("topright", pch=c(17, 15, 16, 3),
 #     col=c("black"),
       legend=c("HBV", "HCV"))
```

the variable "identifiers" is a dataframe that houses all of the metadata information of the samples. The information is pull from the voom output and in order to ensure the sample order matches the expression data we are plotting I use the identical function. The data frame includes the following information (note that many of the number levels seem to be off by one that is intentional as variables were zero indexed and R changed them because it one indexed)

-- The sample ID. Column name is id 

-- The lesion type where 1 means tumor - adjacent and 2 means tumor (identifiers is quickly subsetted to just include tumor samples so all values should say 2). Column name is tumor.

-- Tumor stage where 1 means stage 1, 2 means stage, 3 means stage 3, 4 means stage 4. Column name is stage.

-- Sex of the sample where M means male and F means female. Column name is sex.

-- Edmondson grade of the tumor sample where 1 means grade 1, 1~2 means grade 1.5, 2 means grade 2, 2~3 means grade 2.5, 3 means grade 3 and 4 means grade 4. NA values were replaced with zeros because we are assuming the lack of a value means that the tumor was not graded. Column name is grade 

-- Viral etiology of the tumor sample where 1 means HBV and 2 means HCV. Column name is viral.

-- Whether or not patient smokes where 1 means sample does not smoke and two means sample does smoke. Column name is smoking

-- Level of liver fibrosis in the sample. Where 1 means liver fibrosis severity of zero, 2 means liver fibrosis severity of one, 3 means liver fibrosis severity of two, 4 means liver fibrosis severity of three, and five means liver fibrosis severity of four. Column name is fibrosis.

-- Alcohol intake of the sample where where 1 means alcohol intake level zero, 2 means alcohol intake level one, 3 means alcohol intake level two, 4 means alcohol intake level three. Column name is alcohol

```{r Tumor Samples}
#creating data frame with metadata variables
identifiers<- cbind(id = v[["targets"]][["ID"]], tumor = v[["targets"]][["tumor"]], stage = v[["targets"]][["Ta"]], sex=v[["targets"]][["sex"]], grade = v[["targets"]][["edmonson_grade"]], viral = v[["targets"]][["viral"]], smoking = v[["targets"]][["smoking"]], fibrosis = v[["targets"]][["fibrosis"]], alcohol = v[["targets"]][["alcohol"]])

identifiers<- as.data.frame(identifiers)

#removing NAs fram edmondson grade and replacing with zero
identifiers$grade <- ifelse(is.na(identifiers$grade), 0, identifiers$grade)

#subsetting by tumor samples
subset_condition <- which(identifiers$tumor == "2")
identifiers <- identifiers[subset_condition, ]

#subsetting expression data to tumor samples
Tumor<- vCorrectLibtype[, grepl("tumor",colnames(vCorrectLibtype),fixed=TRUE)]
Tumor<- as.data.frame(Tumor)

#extracting id from sample name since the column name includes additional information
first_five_chars <- substr(colnames(Tumor), 1, 5)

#checking if sample orders match 
result <- identical(identifiers$id, first_five_chars)
print(result)

```

MDS plot was created to see if tumor stage was driving signal. Principal Components 1-4 were checked on the top 25,50,100,1000 genes were checked. To limit the risk of making a mistake the arguments dim.plot was changed from c(1,2) to c(3,4) to check Principal Components 3 and 4 and then top was changed from 25 to 50, 100, and finally 1000 to check the different gene amounts. No clustering was observed.
```{r tumor sample variation}
#creating MDS plot top 25 genes PC 1 and 2 coloring by tumor stage
mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$stage=="1","#FFC20A",
                           ifelse(identifiers$stage=="2", "#E66100",
                                 ifelse(identifiers$stage=="3", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("Stage 1","Stage 2", "Stage 3","Stage 4"))
legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  

```

MDS plot was created to see if viral etiology was driving signal.Principal Components 1-4 were checked on the top 25,50,100,1000 genes were checked. To limit the risk of making a mistake the arguments dim.plot was changed from c(1,2) to c(3,4) to check Principal Components 3 and 4 and then top was changed from 25 to 50, 100, and finally 1000 to check the different gene amounts. No clustering was observed.
```{r virus}
#creating MDS plot top 25 genes principal component 1 and 2 coloring by viral etiology
plotMDS(Tumor, top = 25, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col=ifelse(identifiers$viral %in% c("1"), "#1A85FF", #1 means HBV
                          ifelse(identifiers$viral %in% c("2"), "#D41159", "black")), #2 means HCV
               gene.selection = "common")
legend("top", pch=c(15), 
       col=c("#D41159", "#1A85FF") ,
       legend=c("HCV", "HBV"))
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
```

MDS plot was created to see if Edmondson Grade was driving signal.Principal Components 1-4 were checked on the top 25,50,100,1000 genes were checked. To limit the risk of making a mistake the arguments dim.plot was changed from c(1,2) to c(3,4) to check Principal Components 3 and 4 and then top was changed from 25 to 50, 100, and finally 1000 to check the different gene amounts. No clustering was observed.
```{r edmondson grade}
plotMDS(Tumor, top = 25, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
              col= ifelse(identifiers$grade=="0","red",
                           ifelse(identifiers$grade=="1", "orange",
                                 ifelse(identifiers$grade=="1~2", "yellow",
                                      ifelse(identifiers$grade=="2",  "green",
                                             ifelse(identifiers$grade=="2~3", "darkgreen",
                                                    ifelse(identifiers$grade=="3", "blue", "violet")))))),
               gene.selection = "common")
legend("top", pch=c(15), 
       col=c( "red", "orange","yellow", "green", "darkgreen", "blue", "violet"),
      legend=c("Grade 0","Grade 1", "Grade 1.5","Grade 2", "Grade 2.5", "Grade 3", "Grade 4")) 
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
```

MDS plot was created to see if Smoking was driving signal.Principal Components 1-4 were checked on the top 25,50,100,1000 genes were checked. To limit the risk of making a mistake the arguments dim.plot was changed from c(1,2) to c(3,4) to check Principal Components 3 and 4 and then top was changed from 25 to 50, 100, and finally 1000 to check the different gene amounts. No clustering was observed.
```{r plot smoking}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/smoking.pdf")
plotMDS(Tumor, top = 25, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col=ifelse(identifiers$smoking %in% c("1"), "#D41159", "#1A85FF"),
               gene.selection = "common")
legend("bottomright", pch=c(15), 
       col=c("#1A85FF", "#D41159") ,
       legend=c("Smoking", "Not smoking"))
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
dev.off()

plotMDS(Tumor, top = 25, dim.plot = c(1,2), plot=TRUE, cex=2,
               pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col=ifelse(identifiers$smoking %in% c("1"), "#D41159","#1A85FF"),
               gene.selection = "common")
legend("topleft", pch=c(15), 
       col=c("#1A85FF", "#D41159") ,
       legend=c("Smoking", "Not Smoking"))
legend("center", pch=c(17, 15),
       col=c("black"),
       legend=c("Female", "Male"))
```

MDS plot was created to see if Alcohol Intake was driving signal.Principal Components 1-4 were checked on the top 25,50,100,1000 genes were checked. To limit the risk of making a mistake the arguments dim.plot was changed from c(1,2) to c(3,4) to check Principal Components 3 and 4 and then top was changed from 25 to 50, 100, and finally 1000 to check the different gene amounts. No clustering was observed.
```{r alcohol mds plot}
pdf("~/Desktop/ResearchProjects/LiverCancer/DEA_removed_samples/newFigures/MDSplots/alcohol.pdf")
mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$alcohol=="1","#FFC20A", #one means intake level zero
                           ifelse(identifiers$alcohol=="2", "#E66100", #two means intake level one
                                 ifelse(identifiers$alcohol=="3", "#40B0A6","#006CD1"))), gene.selection = "common") # three means intake level two

legend("bottomleft", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("No alcohol","Severity 1", "Severity 2","Severity 3")) 

legend("topleft", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male")) 
dev.off()

mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$alcohol=="1","#FFC20A",
                           ifelse(identifiers$alcohol=="2", "#E66100",
                                 ifelse(identifiers$alcohol=="3", "#40B0A6","#006CD1"))), gene.selection = "common")

legend("bottom", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1"),
      legend=c("No alcohol","Severity 1", "Severity 2","Severity 3")) 

legend("top", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  
```

MDS plot was created to see if Liver Fibrosis was driving signal.Principal Components 1-4 were checked on the top 25,50,100,1000 genes were checked. To limit the risk of making a mistake the arguments dim.plot was changed from c(1,2) to c(3,4) to check Principal Components 3 and 4 and then top was changed from 25 to 50, 100, and finally 1000 to check the different gene amounts. No clustering was observed.
```{r plot fibrosis}
mds <- plotMDS(Tumor, top = 25, ndim = 10, dim.plot = c(1,2), plot=TRUE, cex=2, 
            pch=ifelse(identifiers$sex %in% c("F"),17, 15),
               col= ifelse(identifiers$fibrosis=="1","#FFC20A", #one = fibrosis level zero
                           ifelse(identifiers$fibrosis=="2", "#E66100", #two = fibrosis level one
                                 ifelse(identifiers$fibrosis=="3", "#40B0A6",#three = fibrosis level two
                                        ifelse(identifiers$fibrosis== "4", "#006CD1", "darkblue")))), gene.selection = "common")

legend("center", pch=c(15), 
      col=c( "#FFC20A", "#E66100","#40B0A6", "#006CD1", "darkblue"),
      legend=c("No Fibrosis","Severity 1", "Severity 2","Severity 3", "Severity 4")) 

legend("topleft", pch=c(17, 15),
     col=c("black"),
legend=c("Female", "Male"))  
```