---
title: "WCNA Analysis HCC data"
author: "Annika Jorgensen"
date: '2022-11-03'
output: html_document
---

## **README**

**Purpose:** To perform the WGCNA analysis on the ICGC Hepatocellular Carcinoma data 

**Methods:** Following the tutorial from https://bioinformaticsworkbook.org/tutorials/wgcna.html#gsc.tab=0 to perform this analysis. Will note decisions and assumptions as I go through in order to doublecheck them with the team. 

**Libraries**
```{r libraries, include=FALSE}
library(tidyverse)
library(magrittr)
library(impute)
library(preprocessCore)
library(WGCNA)
library(plyr)
library(limma)
library(edgeR)
library(biomaRt)
library(org.Mm.eg.db)
library(DGEobj)
library(DGEobj.utils)
```

## **Reading in Data**

```{r read in metadata}
setwd("~/R/Liver Cancer Project")

metadata <- read.table("~/R/Liver Cancer Project/Metadata/metadata_for_de.csv", row.names=1,header=TRUE, sep=",")
tumorAdjacentExp <- read.table("~/R/Liver Cancer Project/Metadata/japan_all_samples_salmon_expression_counts.txt", row.names = 1, header=TRUE)
colnames(tumorAdjacentExp) <- gsub("\\.", "-", colnames(tumorAdjacentExp))

# Importing gene annotations
#genes <- read.table("gencode.v25.chr_patch_hapl_scaff.annotation.bed", header=FALSE, sep="\t")
genes <- read.table("~/R/Liver Cancer Project/Metadata//gencodeTranscripts.txt", header=TRUE, sep="\t")
genes <- data.frame(genes)
tumorAdjacentExp <- tumorAdjacentExp[rownames(tumorAdjacentExp) %in% genes$GENEID ,]
genes <- genes[match(rownames(tumorAdjacentExp), genes$GENEID),]
# Calculating gene length, this is needed for calculating the FPKM values
genes$length <- with(genes, end - start)

#tumorAdjacentExp <- t(tumorAdjacentExp)
#tumorAdjacentExp <- data.frame(tumorAdjacentExpSubset)

# Removing RK023 due to low quality
metadata <- metadata[!(metadata$ID == "RK023") , ]

# Subsetting and ordering metadata to match the count matrix
tumorAdjacentExpSubset <- tumorAdjacentExp[,colnames(tumorAdjacentExp) %in% metadata$sampleid]
#dim(tumorAdjacentExpSubset)
metadataSubset <- metadata[metadata$sampleid %in% colnames(tumorAdjacentExpSubset),]
metadataSubset <- metadataSubset[match(colnames(tumorAdjacentExpSubset), metadataSubset$sampleid),]
rownames(metadataSubset) <- metadataSubset$sampleid
identical(colnames(tumorAdjacentExpSubset),metadataSubset$sampleid)

#tumorAdjacentExpSubset <- t(tumorAdjacentExpSubset)
#tumorAdjacentExpSubset <- data.frame(tumorAdjacentExpSubset)
```

```{r tissue and DGE object}
# Adding tissue type, converting categorical variables to factors
metadataSubset$tumor <- as.numeric(grepl('tumor', metadataSubset$sampleid, ignore.case=T))
metadataSubset$gender_tissue <- paste(metadataSubset$Gender, metadataSubset$tumor, sep="_")
metadataSubset$gender_tissue_viral <- paste(metadataSubset$gender_tissue, metadataSubset$Virus_infection, sep="_")
metadataSubset$library_type <- metadataSubset$strandedness

# Creating the DGEList object

dge <- DGEList(counts=tumorAdjacentExpSubset, genes=genes)

dge$counts <- convertCounts(dge$counts, unit="TPM", geneLength=genes$length)

colnames(dge) <- colnames(tumorAdjacentExpSubset)
dge$samples$sex <- metadataSubset$Gender
dge$samples$viral <- factor(metadataSubset$Virus_infection)
dge$samples$ID <- metadataSubset$ID
dge$samples$tumor <- metadataSubset$tumor
dge$samples$gender_tissue <- metadataSubset$gender_tissue
dge$samples$gender_tissue_viral <- metadataSubset$gender_tissue_viral
dge$samples$library_type <- metadataSubset$library_type
keep <- rowSums(dge$counts>=10) >= 1
dge <- dge[keep,,keep.lib.size=FALSE]


table(dge$samples$gender_tissue_viral)
dim(dge$genes)

TPM<- t(dge$counts)
TPM<-as.data.frame(TPM)
```

```{r threshold calculations, include=FALSE}
allowWGCNAThreads() # Allowing multi-threading

powers = c(c(1:10), seq(from = 5, to = 20, by = 2)) # Choosing a set of soft-thresholding powers, these were selected based on the tutorial only

sft = pickSoftThreshold(TPM, blockSize = 100, powerVector = powers, verbose = 5) # Check with Ken on each of these parameters
```

```{r plot scale independence}
pdf("~/R/Liver Cancer Project/Figures/WGCNA_Scale_Independence.pdf")
plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence")
)
text(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     labels = powers, cex = 0.9, col = "red"
)
abline(h = 0.90, col = "red")
```

```{r Scale Independence}
plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence")
)
text(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     labels = powers, cex = 0.9, col = "red"
)
abline(h = 0.90, col = "red")
```

```{r scale connectivity}
pdf("~/R/Liver Cancer Project/Figures/WGCNA_Scale_Connectivity.pdf")
plot(sft$fitIndices[, 1],
     sft$fitIndices[, 5],
     xlab = "Soft Threshold (power)",
     ylab = "Mean Connectivity",
     type = "n",
     main = paste("Mean connectivity")
)
text(sft$fitIndices[, 1],
     sft$fitIndices[, 5],
     labels = powers,
     cex = 0.9, col = "red")
k=softConnectivity(datExpr = TPM, power=10)
```

```{r print scale connectivity}
plot(sft$fitIndices[, 1],
     sft$fitIndices[, 5],
     xlab = "Soft Threshold (power)",
     ylab = "Mean Connectivity",
     type = "n",
     main = paste("Mean connectivity")
)
text(sft$fitIndices[, 1],
     sft$fitIndices[, 5],
     labels = powers,
     cex = 0.9, col = "red")
k=softConnectivity(datExpr = TPM, power=10)
```

```{r topology}
pdf("~/R/Liver Cancer Project/Figures/WGCNA_scale_free.pdf")
scaleFreePlot(k, main="Check scale free topology\n")
```

```{r print topology}
scaleFreePlot(k, main="Check scale free topology\n")
```

## Calculate and plot WGCNA network

Picking power threshold of 20 based on Mean Connectivity graph and Scale independence graph. Check with Elizabeth just in case 
```{r Calculating WGCNA}
allowWGCNAThreads()
picked_power = 10 
temp_cor <- cor
cor <- WGCNA::cor


netwk <- blockwiseModules(TPM,                # <= input matrix

                          # == Adjacency Function ==
                          power = picked_power,                # <= power here
                          networkType = "signed",

                          # == Tree and Block Options ==
                          deepSplit = 2,
                          pamRespectsDendro = F,
                          # detectCutHeight = 0.75,
                          #minModuleSize = 30,
                          #maxBlockSize = 4000,

                          # == Module Adjustments ==
                          reassignThreshold = 0,
                          mergeCutHeight = 0.25,

                          # == TOM == Archive the run results in TOM file (saves time)
                          saveTOMs = T,
                          saveTOMFileBase = "ER",

                          # == Output Options
                          numericLabels = T,
                          verbose = 3)

# Convert labels to colors for plotting
mergedColors = labels2colors(netwk$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(
  netwk$dendrograms[[1]],
  mergedColors[netwk$blockGenes[[1]]],
  "Modules",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )

pdf("~/R/Liver Cancer Project/Figures/Dendrogram.pdf")
plotDendroAndColors(
  netwk$dendrograms[[1]],
  mergedColors[netwk$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )
dev.off()

cor <- temp_cor
```

## Calculate and plot module-trait correlations
```{r}
module_df <- data.frame(
  gene_id = names(netwk$colors),
  colors = labels2colors(netwk$colors)
)

# Get Module Eigengenes per cluster
MEs0 <- moduleEigengenes(TPM, mergedColors)$eigengenes
print(colnames(MEs0))

module_black <- module_df[which(module_df$colors=="black"),]
black_len <- as.numeric(length(module_black[,1]))
module_blue <- module_df[which(module_df$colors=="blue"),]
blue_len <- as.numeric(length(module_blue[,1]))
module_brown <- module_df[which(module_df$colors=="brown"),]
brown_len <- as.numeric(length(module_brown[,1]))
module_cyan <- module_df[which(module_df$colors=="cyan"),]
cyan_len <- as.numeric(length(module_cyan[,1]))
module_green <- module_df[which(module_df$colors=="green"),]
green_len <- as.numeric(length(module_green[,1]))
module_greenyellow <- module_df[which(module_df$colors=="greenyellow"),]
greenyellow_len <- as.numeric(length(module_greenyellow[,1]))
module_grey <- module_df[which(module_df$colors=="grey"),]
grey_len <- as.numeric(length(module_grey[,1]))
module_lightcyan <- module_df[which(module_df$colors=="lightcyan"),]
lightcyan_len <- as.numeric(length(module_lightcyan[,1]))
module_magenta <- module_df[which(module_df$colors=="magenta"),]
magenta_len <- as.numeric(length(module_magenta[,1]))
module_midnightblue <- module_df[which(module_df$colors=="midnightblue"),]
midnightblue_len <- as.numeric(length(module_midnightblue[,1]))
module_pink <- module_df[which(module_df$colors=="pink"),]
pink_len <- as.numeric(length(module_pink[,1]))
module_purple <- module_df[which(module_df$colors=="purple"),]
purple_len <- as.numeric(length(module_purple[,1]))
module_red <- module_df[which(module_df$colors=="red"),]
red_len <- as.numeric(length(module_red[,1]))
module_salmon <- module_df[which(module_df$colors=="salmon"),]
salmon_len <- as.numeric(length(module_salmon[,1]))
module_tan <- module_df[which(module_df$colors=="tan"),]
tan_len <- as.numeric(length(module_tan[,1]))
module_turquoise <- module_df[which(module_df$colors=="turquoise"),]
turquoise_len <- as.numeric(length(module_turquoise[,1]))
module_yellow <- module_df[which(module_df$colors=="yellow"),]
yellow_len <- as.numeric(length(module_yellow[,1]))

colnames(MEs0) <- c(paste("black (", black_len, ")", sep=""),paste("blue (", blue_len, ")", sep=""),paste("brown (", brown_len, ")",sep=""),
                    paste("cyan (", cyan_len, ")",sep=""), paste("green (", green_len, ")",sep=""), paste("greenyellow (", greenyellow_len, ")",sep=""), 
                    paste("grey (", grey_len, ")", sep=""), paste("lightcyan (", lightcyan_len, ")",sep=""), paste("magenta (", magenta_len, ")",sep=""),
                    paste("midnightblue (", midnightblue_len, ")",sep=""), paste("pink (", pink_len, ")",sep=""), paste("purple (", purple_len, ")",sep=""),
                    paste("red (", red_len, ")", sep=""), paste("salmon (", salmon_len, ")",sep=""), paste("tan (", tan_len, ")",sep=""),
                    paste("turquoise (", turquoise_len, ")", sep=""),paste("yellow (", yellow_len, ")", sep=""))

MEs0 <- MEs0[,17:1]
module_order = names(MEs0) %>% gsub("ME","", .)

# Add treatment names
MEs0$treatment = metadataSubset$gender_tissue_viral

# tidy & plot data
mME = MEs0 %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order)
  )


mME %>% ggplot(., aes(x=fct_inorder(treatment), y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")
```


```{r output module trait}
pdf("~/R/Liver Cancer Project/Figures/Module-trait-relationship.pdf")
mME %>% ggplot(., aes(x=fct_inorder(treatment), y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")
dev.off()
```
